---
title: "A space–time statistical climate model for hurricane intensification in the North Atlantic basin"
author:
    - name: Erik Fraza
      affiliation: Department of Geography, Florida State University
    - name: James B. Elsner
      affiliation: Department of Geography, Florida State University
    - name: Thomas H. Jagger
      affiliation: Department of Geography, Florida State University
date: 2016-08-02
doi: 10.5194/ascmo-2-105-2016
abstract: "Climate influences on hurricane intensification are investigated by averaging hourly intensification rates over the period 1975–2014 in 8°x 8° latitude–longitude grid cells. The statistical effects of hurricane intensity and sea-surface temperature (SST), along with the climatic effects of El Niño–Southern Oscillation (ENSO), the North Atlantic Oscillation (NAO) and the Madden–Julian Oscillation (MJO), are quantified using a Bayesian hierarchical model fit to the averaged data. As expected, stronger hurricanes tend to have higher intensification rates, especially over the warmest waters. Of the three climate variables considered, the NAO has the largest effect on intensification rates after controlling for intensity and SST. The model shows an average increase in intensification rates of 0.18 [0.06, 0.31]ms^-1 h^-1 (95% credible interval) for every 1 standard deviation decrease in the NAO index. Weak trade winds associated with the negative phase of the NAO might result in less vertical wind shear and thus higher mean intensification rates."
output: html_document
---

```{r, include=FALSE}
if (!require("rasterVis")) install.packages("rasterVis", repos = "http://cloud.r-project.org")
if (!require("plotrix")) install.packages("plotrix", repos = "http://cloud.r-project.org")
if (!require("wesanderson")) install.packages("wesanderson", repos = "http://cloud.r-project.org")
```

```{r, include=FALSE}
list.files(pattern="mjomeiAnalysis.Rmd")
library(ggplot2)
library("dplyr")
library("latticeExtra")
load("mjoBestUse.RData")
Intensity = 33.0
Tracks.df = filter(Tracks.df, Int >= Intensity) 

```

```{r, include=FALSE}

fullData.df = load("best.use.2014.Rdata")
begin = 1975; end = 2014

fullData.df = filter(best.use,Yr >= begin, Yr <= end, Int >= 33.0, M == FALSE) 
fullData.df = mutate(best.use,
              Int = WmaxS * .5144,DIntDt = DWmaxDt * .5144) 



```
```{r, include=FALSE}
library("rgdal")
ll = "+proj=longlat +ellps=WGS84"

Tracks.sdf = Tracks.df
coordinates(Tracks.sdf) = c("lon", "lat")
proj4string(Tracks.sdf) = CRS(ll)

fullData.sdf = fullData.df
coordinates(fullData.sdf) = c("lon", "lat")
proj4string(fullData.sdf) = CRS(ll)
```


```{r, include=FALSE}
library("raster")
r = raster(ncol = 10, nrow = 5, 
           xmn = -100, xmx = -20, 
           ymn = 10, ymx = 50)

Tracks.grid = rasterize(Tracks.sdf, r,
                        field = 'DIntDt',
                        fun = mean)

test = rasterToPolygons(Tracks.grid)
sid1=1675
Example.sdf = subset(Tracks.sdf, 
                     Sid == sid1 | 
                     Sid == 1677 |
                     Sid == 1683)
Example.grid = rasterize(Example.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)
```

```{r, include=FALSE}
library("RColorBrewer")
library("rasterVis")
range(values(Example.grid), na.rm = TRUE)
rng = seq(0, 1.5, 1.5)
breaks = c(0, 0.25, 0.50, 0.75, 1.0, 1.25, 1.50)
cr = brewer.pal(9, "Purples")
cr = cr[-(1:3)]
vals = levelplot(Example.grid, margin = FALSE, 
          xlab = NULL, ylab = NULL, 
          col.regions = cr, at = rng, 
          colorkey = NULL,
          border = "white", border.lwd = 2, pretty=TRUE,
          par.settings = list(fontsize = list(text = 15)))
```

Figure 1. Tracks and corresponding cells for three 2008 hurricanes: Gustav, Ike, and Omar. Each colored cell indicates at least one storm intensifies within its boundaries.
```{r, include=FALSE}
library(mapproj)
library(maptools)

outlines = as.data.frame(map("world", xlim = c(-100, -20), ylim = c(10, 50), plot = FALSE)[c("x", "y")],color = "gray")

map = geom_path(aes(x, y), inherit.aes = FALSE, data = outlines, alpha = .8, show_guide = FALSE, color = "blue")

ext = as.vector(extent(r))
boundaries = map("world", fill = TRUE, xlim = ext[1:2], 
                 ylim = ext[3:4], plot = FALSE)

IDs = sapply(strsplit(boundaries$names, ":"), function(x) x[1])

bPols <<- map2SpatialPolygons(boundaries, IDs = IDs,
                              proj4string = CRS(projection(r)))
```

```{r, message=FALSE, warning=FALSE, echo=FALSE, out.width=700}
#This chunk produces Figure 1 from the article

require(gridExtra)
require(sp)

PlotFigure1 = function(Tracks.df,vals){
Int2a.df = filter(Tracks.df, Sid == sid1)
Int2b.df = filter(Tracks.df, Sid == 1677)
Int2c.df = filter(Tracks.df, Sid == 1683)
holdera = data.frame(Int2a.df$lon, Int2a.df$lat)
holderb = data.frame(Int2b.df$lon, Int2b.df$lat)
holderc = data.frame(Int2c.df$lon, Int2c.df$lat)
linesa <<- SpatialLines(list(Lines(list(Line(holdera)),
                                 ID = 'Int2a.df$Sid')))
linesb <<- SpatialLines(list(Lines(list(Line(holderb)),
                                 ID = 'Int2b.df$Sid')))
linesc <<- SpatialLines(list(Lines(list(Line(holderc)),
                                 ID = 'Int2c.df$Sid')))
tracking = vals + 
  latticeExtra::layer(sp.polygons(bPols, col = gray(.8))) +
  latticeExtra::layer(sp.lines(linesa, lwd = 2, col = gray(.4))) + 
  latticeExtra::layer(sp.lines(linesb, lwd = 2, col = gray(.4))) + 
  latticeExtra::layer(sp.lines(linesc, lwd = 2, col = gray(.4))) +
  latticeExtra::layer(panel.text(-63.5, 27.5, 'Gustav', col = gray(.4))) +
  latticeExtra::layer(panel.text(-51, 16.5, 'Ike', col = gray(.4))) + 
  latticeExtra::layer(panel.text(-81.5, 16.25, 'Omar', col = gray(.4)))

plot(tracking)
}
PlotFigure1(Tracks.df,vals)
```
```{r, echo = FALSE}

#This chunk handles mean intensification

data1975 = filter(Tracks.df, Yr == 1975)
data1975.sdf = data1975
coordinates(data1975.sdf) = c("lon", "lat")
proj4string(data1975.sdf) = CRS(ll)

data1975.grid = rasterize(data1975.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1976 = filter(Tracks.df, Yr == 1976)
data1976.sdf = data1976
coordinates(data1976.sdf) = c("lon", "lat")
proj4string(data1976.sdf) = CRS(ll)

data1976.grid = rasterize(data1976.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1977 = filter(Tracks.df, Yr == 1977)
data1977.sdf = data1977
coordinates(data1977.sdf) = c("lon", "lat")
proj4string(data1977.sdf) = CRS(ll)

data1977.grid = rasterize(data1977.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1978 = filter(Tracks.df, Yr == 1978)
data1978.sdf = data1978
coordinates(data1978.sdf) = c("lon", "lat")
proj4string(data1978.sdf) = CRS(ll)

data1978.grid = rasterize(data1978.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1979 = filter(Tracks.df, Yr == 1979)
data1979.sdf = data1979
coordinates(data1979.sdf) = c("lon", "lat")
proj4string(data1979.sdf) = CRS(ll)

data1979.grid = rasterize(data1979.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1980 = filter(Tracks.df, Yr == 1980)
data1980.sdf = data1980
coordinates(data1980.sdf) = c("lon", "lat")
proj4string(data1980.sdf) = CRS(ll)

data1980.grid = rasterize(data1980.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1981 = filter(Tracks.df, Yr == 1981)
data1981.sdf = data1981
coordinates(data1981.sdf) = c("lon", "lat")
proj4string(data1981.sdf) = CRS(ll)

data1981.grid = rasterize(data1981.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1982 = filter(Tracks.df, Yr == 1982)
data1982.sdf = data1982
coordinates(data1982.sdf) = c("lon", "lat")
proj4string(data1982.sdf) = CRS(ll)

data1982.grid = rasterize(data1982.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1983 = filter(Tracks.df, Yr == 1983)
data1983.sdf = data1983
coordinates(data1983.sdf) = c("lon", "lat")
proj4string(data1983.sdf) = CRS(ll)

data1983.grid = rasterize(data1983.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1984 = filter(Tracks.df, Yr == 1984)
data1984.sdf = data1984
coordinates(data1984.sdf) = c("lon", "lat")
proj4string(data1984.sdf) = CRS(ll)

data1984.grid = rasterize(data1984.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1985 = filter(Tracks.df, Yr == 1985)
data1985.sdf = data1985
coordinates(data1985.sdf) = c("lon", "lat")
proj4string(data1985.sdf) = CRS(ll)

data1985.grid = rasterize(data1985.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1986 = filter(Tracks.df, Yr == 1986)
data1986.sdf = data1986
coordinates(data1986.sdf) = c("lon", "lat")
proj4string(data1986.sdf) = CRS(ll)

data1986.grid = rasterize(data1986.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1987 = filter(Tracks.df, Yr == 1987)
data1987.sdf = data1987
coordinates(data1987.sdf) = c("lon", "lat")
proj4string(data1987.sdf) = CRS(ll)

data1987.grid = rasterize(data1987.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1988 = filter(Tracks.df, Yr == 1988)
data1988.sdf = data1988
coordinates(data1988.sdf) = c("lon", "lat")
proj4string(data1988.sdf) = CRS(ll)

data1988.grid = rasterize(data1988.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1989 = filter(Tracks.df, Yr == 1989)
data1989.sdf = data1989
coordinates(data1989.sdf) = c("lon", "lat")
proj4string(data1989.sdf) = CRS(ll)

data1989.grid = rasterize(data1989.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1990 = filter(Tracks.df, Yr == 1990)
data1990.sdf = data1990
coordinates(data1990.sdf) = c("lon", "lat")
proj4string(data1990.sdf) = CRS(ll)

data1990.grid = rasterize(data1990.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1991 = filter(Tracks.df, Yr == 1991)
data1991.sdf = data1991
coordinates(data1991.sdf) = c("lon", "lat")
proj4string(data1991.sdf) = CRS(ll)

data1991.grid = rasterize(data1991.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1992 = filter(Tracks.df, Yr == 1992)
data1992.sdf = data1992
coordinates(data1992.sdf) = c("lon", "lat")
proj4string(data1992.sdf) = CRS(ll)

data1992.grid = rasterize(data1992.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1993 = filter(Tracks.df, Yr == 1993)
data1993.sdf = data1993
coordinates(data1993.sdf) = c("lon", "lat")
proj4string(data1993.sdf) = CRS(ll)

data1993.grid = rasterize(data1993.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1994 = filter(Tracks.df, Yr == 1994)
data1994.sdf = data1994
coordinates(data1994.sdf) = c("lon", "lat")
proj4string(data1994.sdf) = CRS(ll)

data1994.grid = rasterize(data1994.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1995 = filter(Tracks.df, Yr == 1995)
data1995.sdf = data1995
coordinates(data1995.sdf) = c("lon", "lat")
proj4string(data1995.sdf) = CRS(ll)

data1995.grid = rasterize(data1995.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1996 = filter(Tracks.df, Yr == 1996)
data1996.sdf = data1996
coordinates(data1996.sdf) = c("lon", "lat")
proj4string(data1996.sdf) = CRS(ll)

data1996.grid = rasterize(data1996.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1997 = filter(Tracks.df, Yr == 1997)
data1997.sdf = data1997
coordinates(data1997.sdf) = c("lon", "lat")
proj4string(data1997.sdf) = CRS(ll)

data1997.grid = rasterize(data1997.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1998 = filter(Tracks.df, Yr == 1998)
data1998.sdf = data1998
coordinates(data1998.sdf) = c("lon", "lat")
proj4string(data1998.sdf) = CRS(ll)

data1998.grid = rasterize(data1998.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data1999 = filter(Tracks.df, Yr == 1999)
data1999.sdf = data1999
coordinates(data1999.sdf) = c("lon", "lat")
proj4string(data1999.sdf) = CRS(ll)

data1999.grid = rasterize(data1999.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data2000 = filter(Tracks.df, Yr == 2000)
data2000.sdf = data2000
coordinates(data2000.sdf) = c("lon", "lat")
proj4string(data2000.sdf) = CRS(ll)

data2000.grid = rasterize(data2000.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data2001 = filter(Tracks.df, Yr == 2001)
data2001.sdf = data2001
coordinates(data2001.sdf) = c("lon", "lat")
proj4string(data2001.sdf) = CRS(ll)

data2001.grid = rasterize(data2001.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data2002 = filter(Tracks.df, Yr == 2002)
data2002.sdf = data2002
coordinates(data2002.sdf) = c("lon", "lat")
proj4string(data2002.sdf) = CRS(ll)

data2002.grid = rasterize(data2002.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data2003 = filter(Tracks.df, Yr == 2003)
data2003.sdf = data2003
coordinates(data2003.sdf) = c("lon", "lat")
proj4string(data2003.sdf) = CRS(ll)

data2003.grid = rasterize(data2003.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data2004 = filter(Tracks.df, Yr == 2004)
data2004.sdf = data2004
coordinates(data2004.sdf) = c("lon", "lat")
proj4string(data2004.sdf) = CRS(ll)

data2004.grid = rasterize(data2004.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data2005 = filter(Tracks.df, Yr == 2005)
data2005.sdf = data2005
coordinates(data2005.sdf) = c("lon", "lat")
proj4string(data2005.sdf) = CRS(ll)

data2005.grid = rasterize(data2005.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data2006 = filter(Tracks.df, Yr == 2006)
data2006.sdf = data2006
coordinates(data2006.sdf) = c("lon", "lat")
proj4string(data2006.sdf) = CRS(ll)

data2006.grid = rasterize(data2006.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data2007 = filter(Tracks.df, Yr == 2007)
data2007.sdf = data2007
coordinates(data2007.sdf) = c("lon", "lat")
proj4string(data2007.sdf) = CRS(ll)

data2007.grid = rasterize(data2007.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data2008 = filter(Tracks.df, Yr == 2008)
data2008.sdf = data2008
coordinates(data2008.sdf) = c("lon", "lat")
proj4string(data2008.sdf) = CRS(ll)

data2008.grid = rasterize(data2008.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data2009 = filter(Tracks.df, Yr == 2009)
data2009.sdf = data2009
coordinates(data2009.sdf) = c("lon", "lat")
proj4string(data2009.sdf) = CRS(ll)

data2009.grid = rasterize(data2009.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data2010 = filter(Tracks.df, Yr == 2010)
data2010.sdf = data2010
coordinates(data2010.sdf) = c("lon", "lat")
proj4string(data2010.sdf) = CRS(ll)

data2010.grid = rasterize(data2010.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data2011 = filter(Tracks.df, Yr == 2011)
data2011.sdf = data2011
coordinates(data2011.sdf) = c("lon", "lat")
proj4string(data2011.sdf) = CRS(ll)

data2011.grid = rasterize(data2011.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data2012 = filter(Tracks.df, Yr == 2012)
data2012.sdf = data2012
coordinates(data2012.sdf) = c("lon", "lat")
proj4string(data2012.sdf) = CRS(ll)

data2012.grid = rasterize(data2012.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data2013 = filter(Tracks.df, Yr == 2013)
data2013.sdf = data2013
coordinates(data2013.sdf) = c("lon", "lat")
proj4string(data2013.sdf) = CRS(ll)

data2013.grid = rasterize(data2013.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

data2014 = filter(Tracks.df, Yr == 2014)
data2014.sdf = data2014
coordinates(data2014.sdf) = c("lon", "lat")
proj4string(data2014.sdf) = CRS(ll)

data2014.grid = rasterize(data2014.sdf, r, 
                         field = 'DIntDt',
                         fun = mean)

dataDIntDt1975.df = as.data.frame(data1975.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1975.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1976.df = as.data.frame(data1976.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1976.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1977.df = as.data.frame(data1977.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1977.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1978.df = as.data.frame(data1978.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1978.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1979.df = as.data.frame(data1979.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1979.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1980.df = as.data.frame(data1980.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1980.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1981.df = as.data.frame(data1981.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1981.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1982.df = as.data.frame(data1982.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1982.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1983.df = as.data.frame(data1983.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1983.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1984.df = as.data.frame(data1984.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1984.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1985.df = as.data.frame(data1985.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1985.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1986.df = as.data.frame(data1986.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1986.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1987.df = as.data.frame(data1987.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1987.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1988.df = as.data.frame(data1988.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1988.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1989.df = as.data.frame(data1989.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1989.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1990.df = as.data.frame(data1990.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1990.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1991.df = as.data.frame(data1991.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1991.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1992.df = as.data.frame(data1992.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1992.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1993.df = as.data.frame(data1993.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1993.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1994.df = as.data.frame(data1994.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1994.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1995.df = as.data.frame(data1995.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1995.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1996.df = as.data.frame(data1996.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1996.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1997.df = as.data.frame(data1997.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1997.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1998.df = as.data.frame(data1998.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1998.df) <- c("lon", "lat", "DIntDt")
dataDIntDt1999.df = as.data.frame(data1999.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt1999.df) <- c("lon", "lat", "DIntDt")
dataDIntDt2000.df = as.data.frame(data2000.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt2000.df) <- c("lon", "lat", "DIntDt")
dataDIntDt2001.df = as.data.frame(data2001.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt2001.df) <- c("lon", "lat", "DIntDt")
dataDIntDt2002.df = as.data.frame(data2002.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt2002.df) <- c("lon", "lat", "DIntDt")
dataDIntDt2003.df = as.data.frame(data2003.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt2003.df) <- c("lon", "lat", "DIntDt")
dataDIntDt2004.df = as.data.frame(data2004.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt2004.df) <- c("lon", "lat", "DIntDt")
dataDIntDt2005.df = as.data.frame(data2005.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt2005.df) <- c("lon", "lat", "DIntDt")
dataDIntDt2006.df = as.data.frame(data2006.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt2006.df) <- c("lon", "lat", "DIntDt")
dataDIntDt2007.df = as.data.frame(data2007.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt2007.df) <- c("lon", "lat", "DIntDt")
dataDIntDt2008.df = as.data.frame(data2008.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt2008.df) <- c("lon", "lat", "DIntDt")
dataDIntDt2009.df = as.data.frame(data2009.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt2009.df) <- c("lon", "lat", "DIntDt")
dataDIntDt2010.df = as.data.frame(data2010.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt2010.df) <- c("lon", "lat", "DIntDt")
dataDIntDt2011.df = as.data.frame(data2011.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt2011.df) <- c("lon", "lat", "DIntDt")
dataDIntDt2012.df = as.data.frame(data2012.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt2012.df) <- c("lon", "lat", "DIntDt")
dataDIntDt2013.df = as.data.frame(data2013.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt2013.df) <- c("lon", "lat", "DIntDt")
dataDIntDt2014.df = as.data.frame(data2014.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataDIntDt2014.df) <- c("lon", "lat", "DIntDt")

dataDIntDt1975.df$Year[1:50] = 1975
dataDIntDt1976.df$Year[1:50] = 1976
dataDIntDt1977.df$Year[1:50] = 1977
dataDIntDt1978.df$Year[1:50] = 1978
dataDIntDt1979.df$Year[1:50] = 1979
dataDIntDt1980.df$Year[1:50] = 1980
dataDIntDt1981.df$Year[1:50] = 1981
dataDIntDt1982.df$Year[1:50] = 1982
dataDIntDt1983.df$Year[1:50] = 1983
dataDIntDt1984.df$Year[1:50] = 1984
dataDIntDt1985.df$Year[1:50] = 1985
dataDIntDt1986.df$Year[1:50] = 1986
dataDIntDt1987.df$Year[1:50] = 1987
dataDIntDt1988.df$Year[1:50] = 1988
dataDIntDt1989.df$Year[1:50] = 1989
dataDIntDt1990.df$Year[1:50] = 1990
dataDIntDt1991.df$Year[1:50] = 1991
dataDIntDt1992.df$Year[1:50] = 1992
dataDIntDt1993.df$Year[1:50] = 1993
dataDIntDt1994.df$Year[1:50] = 1994
dataDIntDt1995.df$Year[1:50] = 1995
dataDIntDt1996.df$Year[1:50] = 1996
dataDIntDt1997.df$Year[1:50] = 1997
dataDIntDt1998.df$Year[1:50] = 1998
dataDIntDt1999.df$Year[1:50] = 1999
dataDIntDt2000.df$Year[1:50] = 2000
dataDIntDt2001.df$Year[1:50] = 2001
dataDIntDt2002.df$Year[1:50] = 2002
dataDIntDt2003.df$Year[1:50] = 2003
dataDIntDt2004.df$Year[1:50] = 2004
dataDIntDt2005.df$Year[1:50] = 2005
dataDIntDt2006.df$Year[1:50] = 2006
dataDIntDt2007.df$Year[1:50] = 2007
dataDIntDt2008.df$Year[1:50] = 2008
dataDIntDt2009.df$Year[1:50] = 2009
dataDIntDt2010.df$Year[1:50] = 2010
dataDIntDt2011.df$Year[1:50] = 2011
dataDIntDt2012.df$Year[1:50] = 2012
dataDIntDt2013.df$Year[1:50] = 2013
dataDIntDt2014.df$Year[1:50] = 2014

dataDIntDt1975.df$ID = as.numeric(as.factor(with(dataDIntDt1975.df, paste(lon, lat,  sep="_"))))
dataDIntDt1976.df$ID = as.numeric(as.factor(with(dataDIntDt1976.df, paste(lon, lat,  sep="_"))))
dataDIntDt1977.df$ID = as.numeric(as.factor(with(dataDIntDt1977.df, paste(lon, lat,  sep="_"))))
dataDIntDt1978.df$ID = as.numeric(as.factor(with(dataDIntDt1978.df, paste(lon, lat,  sep="_"))))
dataDIntDt1979.df$ID = as.numeric(as.factor(with(dataDIntDt1979.df, paste(lon, lat,  sep="_"))))
dataDIntDt1980.df$ID = as.numeric(as.factor(with(dataDIntDt1980.df, paste(lon, lat,  sep="_"))))
dataDIntDt1981.df$ID = as.numeric(as.factor(with(dataDIntDt1981.df, paste(lon, lat,  sep="_"))))
dataDIntDt1982.df$ID = as.numeric(as.factor(with(dataDIntDt1982.df, paste(lon, lat,  sep="_"))))
dataDIntDt1983.df$ID = as.numeric(as.factor(with(dataDIntDt1983.df, paste(lon, lat,  sep="_"))))
dataDIntDt1984.df$ID = as.numeric(as.factor(with(dataDIntDt1984.df, paste(lon, lat,  sep="_"))))
dataDIntDt1985.df$ID = as.numeric(as.factor(with(dataDIntDt1985.df, paste(lon, lat,  sep="_"))))
dataDIntDt1986.df$ID = as.numeric(as.factor(with(dataDIntDt1986.df, paste(lon, lat,  sep="_"))))
dataDIntDt1987.df$ID = as.numeric(as.factor(with(dataDIntDt1987.df, paste(lon, lat,  sep="_"))))
dataDIntDt1988.df$ID = as.numeric(as.factor(with(dataDIntDt1988.df, paste(lon, lat,  sep="_"))))
dataDIntDt1989.df$ID = as.numeric(as.factor(with(dataDIntDt1989.df, paste(lon, lat,  sep="_"))))
dataDIntDt1990.df$ID = as.numeric(as.factor(with(dataDIntDt1990.df, paste(lon, lat,  sep="_"))))
dataDIntDt1991.df$ID = as.numeric(as.factor(with(dataDIntDt1991.df, paste(lon, lat,  sep="_"))))
dataDIntDt1992.df$ID = as.numeric(as.factor(with(dataDIntDt1992.df, paste(lon, lat,  sep="_"))))
dataDIntDt1993.df$ID = as.numeric(as.factor(with(dataDIntDt1993.df, paste(lon, lat,  sep="_"))))
dataDIntDt1994.df$ID = as.numeric(as.factor(with(dataDIntDt1994.df, paste(lon, lat,  sep="_"))))
dataDIntDt1995.df$ID = as.numeric(as.factor(with(dataDIntDt1995.df, paste(lon, lat,  sep="_"))))
dataDIntDt1996.df$ID = as.numeric(as.factor(with(dataDIntDt1996.df, paste(lon, lat,  sep="_"))))
dataDIntDt1997.df$ID = as.numeric(as.factor(with(dataDIntDt1997.df, paste(lon, lat,  sep="_"))))
dataDIntDt1998.df$ID = as.numeric(as.factor(with(dataDIntDt1998.df, paste(lon, lat,  sep="_"))))
dataDIntDt1999.df$ID = as.numeric(as.factor(with(dataDIntDt1999.df, paste(lon, lat,  sep="_"))))
dataDIntDt2000.df$ID = as.numeric(as.factor(with(dataDIntDt2000.df, paste(lon, lat,  sep="_"))))
dataDIntDt2001.df$ID = as.numeric(as.factor(with(dataDIntDt2001.df, paste(lon, lat,  sep="_"))))
dataDIntDt2002.df$ID = as.numeric(as.factor(with(dataDIntDt2002.df, paste(lon, lat,  sep="_"))))
dataDIntDt2003.df$ID = as.numeric(as.factor(with(dataDIntDt2003.df, paste(lon, lat,  sep="_"))))
dataDIntDt2004.df$ID = as.numeric(as.factor(with(dataDIntDt2004.df, paste(lon, lat,  sep="_"))))
dataDIntDt2005.df$ID = as.numeric(as.factor(with(dataDIntDt2005.df, paste(lon, lat,  sep="_"))))
dataDIntDt2006.df$ID = as.numeric(as.factor(with(dataDIntDt2006.df, paste(lon, lat,  sep="_"))))
dataDIntDt2007.df$ID = as.numeric(as.factor(with(dataDIntDt2007.df, paste(lon, lat,  sep="_"))))
dataDIntDt2008.df$ID = as.numeric(as.factor(with(dataDIntDt2008.df, paste(lon, lat,  sep="_"))))
dataDIntDt2009.df$ID = as.numeric(as.factor(with(dataDIntDt2009.df, paste(lon, lat,  sep="_"))))
dataDIntDt2010.df$ID = as.numeric(as.factor(with(dataDIntDt2010.df, paste(lon, lat,  sep="_"))))
dataDIntDt2011.df$ID = as.numeric(as.factor(with(dataDIntDt2011.df, paste(lon, lat,  sep="_"))))
dataDIntDt2012.df$ID = as.numeric(as.factor(with(dataDIntDt2012.df, paste(lon, lat,  sep="_"))))
dataDIntDt2013.df$ID = as.numeric(as.factor(with(dataDIntDt2013.df, paste(lon, lat,  sep="_"))))
dataDIntDt2014.df$ID = as.numeric(as.factor(with(dataDIntDt2014.df, paste(lon, lat,  sep="_"))))

```

```{r, echo = FALSE}

Intensification = rbind(dataDIntDt1975.df,dataDIntDt1976.df, dataDIntDt1977.df, dataDIntDt1978.df, dataDIntDt1979.df,
                        dataDIntDt1980.df, dataDIntDt1981.df, dataDIntDt1982.df, dataDIntDt1983.df,
                        dataDIntDt1984.df, dataDIntDt1985.df, dataDIntDt1986.df, dataDIntDt1987.df,
                        dataDIntDt1988.df, dataDIntDt1989.df, dataDIntDt1990.df, dataDIntDt1991.df, 
                        dataDIntDt1992.df, dataDIntDt1993.df, dataDIntDt1994.df, dataDIntDt1995.df, 
                        dataDIntDt1996.df, dataDIntDt1997.df, dataDIntDt1998.df, dataDIntDt1999.df,
                        dataDIntDt2000.df, dataDIntDt2001.df, dataDIntDt2002.df, dataDIntDt2003.df,
                        dataDIntDt2004.df, dataDIntDt2005.df, dataDIntDt2006.df, dataDIntDt2007.df,
                        dataDIntDt2008.df, dataDIntDt2009.df, dataDIntDt2010.df, dataDIntDt2011.df,
                        dataDIntDt2012.df, dataDIntDt2013.df, dataDIntDt2014.df) 

Intensification$ID = as.integer(Intensification$ID)

looper = nrow(Intensification)

while(looper > 0){
  if ((Intensification$ID[looper] == 50) == TRUE | (Intensification$ID[looper] == 46) == TRUE |
        (Intensification$ID[looper] == 45) == TRUE | (Intensification$ID[looper] == 35) == TRUE |
        (Intensification$ID[looper] == 49) == TRUE | (Intensification$ID[looper] == 40) == TRUE|
        (Intensification$ID[looper] == 44) == TRUE | (Intensification$ID[looper] == 3) == TRUE |
        (Intensification$ID[looper] == 4) == TRUE)
    {
      Intensification = Intensification[-looper,]
      }
  looper = looper - 1
  }

```

```{r, echo = FALSE}

ID = dataDIntDt1975.df$ID
Yr = c(1975:2014)

DIntDtData.df = data.frame(ID)

DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1975.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1976.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1977.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1978.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1979.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1980.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1981.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1982.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1983.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1984.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1985.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1986.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1987.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1988.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1989.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1990.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1991.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1992.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1993.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1994.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1995.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1996.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1997.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1998.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt1999.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt2000.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt2001.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt2002.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt2003.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt2004.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt2005.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt2006.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt2007.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt2008.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt2009.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt2010.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt2011.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt2012.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt2013.df$DIntDt)
DIntDtData.df = cbind(DIntDtData.df, dataDIntDt2014.df$DIntDt)

x = 1
while (x < 41){
  colnames(DIntDtData.df)[x+1] = Yr[x]
  x = x+1
}

DIntDtData.df$ID = as.integer(DIntDtData.df$ID)

looper = nrow(DIntDtData.df)

while(looper > 0){
  if ((DIntDtData.df$ID[looper] == 50) == TRUE | (DIntDtData.df$ID[looper] == 46) == TRUE |
        (DIntDtData.df$ID[looper] == 45) == TRUE | (DIntDtData.df$ID[looper] == 35) == TRUE |
        (DIntDtData.df$ID[looper] == 49) == TRUE | (DIntDtData.df$ID[looper] == 40) == TRUE|
        (DIntDtData.df$ID[looper] == 44) == TRUE | (DIntDtData.df$ID[looper] == 3) == TRUE |
        (DIntDtData.df$ID[looper] == 4) == TRUE)
    {
      DIntDtData.df = DIntDtData.df[-looper,]
      }
  looper = looper - 1
}

idOrder = DIntDtData.df$ID
latOrder = Intensification$lat[1:41]
lonOrder = Intensification$lon[1:41]

gridOrder = data.frame(ID = idOrder, lat = latOrder, lon = lonOrder)
gridOrder = arrange(gridOrder, ID)

```

```{r, echo = FALSE}

DIntDtStack = stack(data1975.grid, data1976.grid, data1977.grid, data1978.grid, data1979.grid, data1980.grid,
                    data1981.grid, data1982.grid, data1983.grid, data1984.grid, data1985.grid,
                    data1986.grid, data1987.grid, data1988.grid, data1989.grid, data1990.grid,
                    data1991.grid, data1992.grid, data1993.grid, data1994.grid, data1995.grid,
                    data1996.grid, data1997.grid, data1998.grid, data1999.grid, data2000.grid,
                    data2001.grid, data2002.grid, data2003.grid, data2004.grid, data2005.grid,
                    data2006.grid, data2007.grid, data2008.grid, data2009.grid, data2010.grid,
                    data2011.grid, data2012.grid, data2013.grid, data2014.grid)

DIntDtStackCount = nlayers(DIntDtStack)

```

```{r, echo = FALSE}
#This chunk produces Figure 4 from the article.
library("latticeExtra")

library("rasterVis")

names = c('1975', '1976', '1977', '1978', '1979', '1980', '1981', '1982', '1983', '1984','1985', '1986', '1987', '1988', '1989', '1990', '1991', '1992', '1993',
          '1994', '1995', '1996', '1997', '1998', '1999', '2000', '2001', '2002',
          '2003', '2004', '2005', '2006', '2007', '2008', '2009', '2010', '2011',
          '2012', '2013', '2014')

background = sp.polygons(bPols, col = gray(.8))

cr = brewer.pal(9, "Purples")
cr = cr[-(1)]
rng = seq(0, 2.00, 0.25)
breaks = c(0.00, 0.25, 0.50, 0.75, 1.00, 1.25, 1.50, 1.75, 2.00 )

plotting = spplot(DIntDtStack, col.regions=cr, names.attr = names,
                  at = rng, colorkey = list(space="bottom", at=breaks, labels=list(at=breaks)), labels=paste(rng), par.settings=list(fontsize=list(text=10)),
        sub=expression(paste("Intensification [m ", s^-1, " per h]")),
              pretty=TRUE)

p1 = plotting + latticeExtra::layer(sp.polygons(bPols))
plot(p1)

```

```{r, echo = FALSE}

#This chunk handles mean intensity

data1975 = filter(Tracks.df, Yr == 1975)
data1975.sdf = data1975
coordinates(data1975.sdf) = c("lon", "lat")
proj4string(data1975.sdf) = CRS(ll)

data1975.grid = rasterize(data1975.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1976 = filter(Tracks.df, Yr == 1976)
data1976.sdf = data1976
coordinates(data1976.sdf) = c("lon", "lat")
proj4string(data1976.sdf) = CRS(ll)

data1976.grid = rasterize(data1976.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1977 = filter(Tracks.df, Yr == 1977)
data1977.sdf = data1977
coordinates(data1977.sdf) = c("lon", "lat")
proj4string(data1977.sdf) = CRS(ll)

data1977.grid = rasterize(data1977.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1978 = filter(Tracks.df, Yr == 1978)
data1978.sdf = data1978
coordinates(data1978.sdf) = c("lon", "lat")
proj4string(data1978.sdf) = CRS(ll)

data1978.grid = rasterize(data1978.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1979 = filter(Tracks.df, Yr == 1979)
data1979.sdf = data1979
coordinates(data1979.sdf) = c("lon", "lat")
proj4string(data1979.sdf) = CRS(ll)

data1979.grid = rasterize(data1979.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1980 = filter(Tracks.df, Yr == 1980)
data1980.sdf = data1980
coordinates(data1980.sdf) = c("lon", "lat")
proj4string(data1980.sdf) = CRS(ll)

data1980.grid = rasterize(data1980.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1981 = filter(Tracks.df, Yr == 1981)
data1981.sdf = data1981
coordinates(data1981.sdf) = c("lon", "lat")
proj4string(data1981.sdf) = CRS(ll)

data1981.grid = rasterize(data1981.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1982 = filter(Tracks.df, Yr == 1982)
data1982.sdf = data1982
coordinates(data1982.sdf) = c("lon", "lat")
proj4string(data1982.sdf) = CRS(ll)

data1982.grid = rasterize(data1982.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1983 = filter(Tracks.df, Yr == 1983)
data1983.sdf = data1983
coordinates(data1983.sdf) = c("lon", "lat")
proj4string(data1983.sdf) = CRS(ll)

data1983.grid = rasterize(data1983.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1984 = filter(Tracks.df, Yr == 1984)
data1984.sdf = data1984
coordinates(data1984.sdf) = c("lon", "lat")
proj4string(data1984.sdf) = CRS(ll)

data1984.grid = rasterize(data1984.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1985 = filter(Tracks.df, Yr == 1985)
data1985.sdf = data1985
coordinates(data1985.sdf) = c("lon", "lat")
proj4string(data1985.sdf) = CRS(ll)

data1985.grid = rasterize(data1985.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1986 = filter(Tracks.df, Yr == 1986)
data1986.sdf = data1986
coordinates(data1986.sdf) = c("lon", "lat")
proj4string(data1986.sdf) = CRS(ll)

data1986.grid = rasterize(data1986.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1987 = filter(Tracks.df, Yr == 1987)
data1987.sdf = data1987
coordinates(data1987.sdf) = c("lon", "lat")
proj4string(data1987.sdf) = CRS(ll)

data1987.grid = rasterize(data1987.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1988 = filter(Tracks.df, Yr == 1988)
data1988.sdf = data1988
coordinates(data1988.sdf) = c("lon", "lat")
proj4string(data1988.sdf) = CRS(ll)

data1988.grid = rasterize(data1988.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1989 = filter(Tracks.df, Yr == 1989)
data1989.sdf = data1989
coordinates(data1989.sdf) = c("lon", "lat")
proj4string(data1989.sdf) = CRS(ll)

data1989.grid = rasterize(data1989.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1990 = filter(Tracks.df, Yr == 1990)
data1990.sdf = data1990
coordinates(data1990.sdf) = c("lon", "lat")
proj4string(data1990.sdf) = CRS(ll)

data1990.grid = rasterize(data1990.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1991 = filter(Tracks.df, Yr == 1991)
data1991.sdf = data1991
coordinates(data1991.sdf) = c("lon", "lat")
proj4string(data1991.sdf) = CRS(ll)

data1991.grid = rasterize(data1991.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1992 = filter(Tracks.df, Yr == 1992)
data1992.sdf = data1992
coordinates(data1992.sdf) = c("lon", "lat")
proj4string(data1992.sdf) = CRS(ll)

data1992.grid = rasterize(data1992.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1993 = filter(Tracks.df, Yr == 1993)
data1993.sdf = data1993
coordinates(data1993.sdf) = c("lon", "lat")
proj4string(data1993.sdf) = CRS(ll)

data1993.grid = rasterize(data1993.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1994 = filter(Tracks.df, Yr == 1994)
data1994.sdf = data1994
coordinates(data1994.sdf) = c("lon", "lat")
proj4string(data1994.sdf) = CRS(ll)

data1994.grid = rasterize(data1994.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1995 = filter(Tracks.df, Yr == 1995)
data1995.sdf = data1995
coordinates(data1995.sdf) = c("lon", "lat")
proj4string(data1995.sdf) = CRS(ll)

data1995.grid = rasterize(data1995.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1996 = filter(Tracks.df, Yr == 1996)
data1996.sdf = data1996
coordinates(data1996.sdf) = c("lon", "lat")
proj4string(data1996.sdf) = CRS(ll)

data1996.grid = rasterize(data1996.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1997 = filter(Tracks.df, Yr == 1997)
data1997.sdf = data1997
coordinates(data1997.sdf) = c("lon", "lat")
proj4string(data1997.sdf) = CRS(ll)

data1997.grid = rasterize(data1997.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1998 = filter(Tracks.df, Yr == 1998)
data1998.sdf = data1998
coordinates(data1998.sdf) = c("lon", "lat")
proj4string(data1998.sdf) = CRS(ll)

data1998.grid = rasterize(data1998.sdf, r, 
                         field = 'Int',
                         fun = mean)

data1999 = filter(Tracks.df, Yr == 1999)
data1999.sdf = data1999
coordinates(data1999.sdf) = c("lon", "lat")
proj4string(data1999.sdf) = CRS(ll)

data1999.grid = rasterize(data1999.sdf, r, 
                         field = 'Int',
                         fun = mean)

data2000 = filter(Tracks.df, Yr == 2000)
data2000.sdf = data2000
coordinates(data2000.sdf) = c("lon", "lat")
proj4string(data2000.sdf) = CRS(ll)

data2000.grid = rasterize(data2000.sdf, r, 
                         field = 'Int',
                         fun = mean)

data2001 = filter(Tracks.df, Yr == 2001)
data2001.sdf = data2001
coordinates(data2001.sdf) = c("lon", "lat")
proj4string(data2001.sdf) = CRS(ll)

data2001.grid = rasterize(data2001.sdf, r, 
                         field = 'Int',
                         fun = mean)

data2002 = filter(Tracks.df, Yr == 2002)
data2002.sdf = data2002
coordinates(data2002.sdf) = c("lon", "lat")
proj4string(data2002.sdf) = CRS(ll)

data2002.grid = rasterize(data2002.sdf, r, 
                         field = 'Int',
                         fun = mean)

data2003 = filter(Tracks.df, Yr == 2003)
data2003.sdf = data2003
coordinates(data2003.sdf) = c("lon", "lat")
proj4string(data2003.sdf) = CRS(ll)

data2003.grid = rasterize(data2003.sdf, r, 
                         field = 'Int',
                         fun = mean)

data2004 = filter(Tracks.df, Yr == 2004)
data2004.sdf = data2004
coordinates(data2004.sdf) = c("lon", "lat")
proj4string(data2004.sdf) = CRS(ll)

data2004.grid = rasterize(data2004.sdf, r, 
                         field = 'Int',
                         fun = mean)

data2005 = filter(Tracks.df, Yr == 2005)
data2005.sdf = data2005
coordinates(data2005.sdf) = c("lon", "lat")
proj4string(data2005.sdf) = CRS(ll)

data2005.grid = rasterize(data2005.sdf, r, 
                         field = 'Int',
                         fun = mean)

data2006 = filter(Tracks.df, Yr == 2006)
data2006.sdf = data2006
coordinates(data2006.sdf) = c("lon", "lat")
proj4string(data2006.sdf) = CRS(ll)

data2006.grid = rasterize(data2006.sdf, r, 
                         field = 'Int',
                         fun = mean)

data2007 = filter(Tracks.df, Yr == 2007)
data2007.sdf = data2007
coordinates(data2007.sdf) = c("lon", "lat")
proj4string(data2007.sdf) = CRS(ll)

data2007.grid = rasterize(data2007.sdf, r, 
                         field = 'Int',
                         fun = mean)

data2008 = filter(Tracks.df, Yr == 2008)
data2008.sdf = data2008
coordinates(data2008.sdf) = c("lon", "lat")
proj4string(data2008.sdf) = CRS(ll)

data2008.grid = rasterize(data2008.sdf, r, 
                         field = 'Int',
                         fun = mean)

data2009 = filter(Tracks.df, Yr == 2009)
data2009.sdf = data2009
coordinates(data2009.sdf) = c("lon", "lat")
proj4string(data2009.sdf) = CRS(ll)

data2009.grid = rasterize(data2009.sdf, r, 
                         field = 'Int',
                         fun = mean)

data2010 = filter(Tracks.df, Yr == 2010)
data2010.sdf = data2010
coordinates(data2010.sdf) = c("lon", "lat")
proj4string(data2010.sdf) = CRS(ll)

data2010.grid = rasterize(data2010.sdf, r, 
                         field = 'Int',
                         fun = mean)

data2011 = filter(Tracks.df, Yr == 2011)
data2011.sdf = data2011
coordinates(data2011.sdf) = c("lon", "lat")
proj4string(data2011.sdf) = CRS(ll)

data2011.grid = rasterize(data2011.sdf, r, 
                         field = 'Int',
                         fun = mean)

data2012 = filter(Tracks.df, Yr == 2012)
data2012.sdf = data2012
coordinates(data2012.sdf) = c("lon", "lat")
proj4string(data2012.sdf) = CRS(ll)

data2012.grid = rasterize(data2012.sdf, r, 
                         field = 'Int',
                         fun = mean)

data2013 = filter(Tracks.df, Yr == 2013)
data2013.sdf = data2013
coordinates(data2013.sdf) = c("lon", "lat")
proj4string(data2013.sdf) = CRS(ll)

data2013.grid = rasterize(data2013.sdf, r, 
                         field = 'Int',
                         fun = mean)

data2014 = filter(Tracks.df, Yr == 2014)
data2014.sdf = data2014
coordinates(data2014.sdf) = c("lon", "lat")
proj4string(data2014.sdf) = CRS(ll)

data2014.grid = rasterize(data2014.sdf, r, 
                         field = 'Int',
                         fun = mean)

dataInt1975.df = as.data.frame(data1975.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1975.df) <- c("lon", "lat", "Int")
dataInt1976.df = as.data.frame(data1976.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1976.df) <- c("lon", "lat", "Int")
dataInt1977.df = as.data.frame(data1977.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1977.df) <- c("lon", "lat", "Int")
dataInt1978.df = as.data.frame(data1978.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1978.df) <- c("lon", "lat", "Int")
dataInt1979.df = as.data.frame(data1979.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1979.df) <- c("lon", "lat", "Int")
dataInt1980.df = as.data.frame(data1980.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1980.df) <- c("lon", "lat", "Int")
dataInt1981.df = as.data.frame(data1981.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1981.df) <- c("lon", "lat", "Int")
dataInt1982.df = as.data.frame(data1982.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1982.df) <- c("lon", "lat", "Int")
dataInt1983.df = as.data.frame(data1983.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1983.df) <- c("lon", "lat", "Int")
dataInt1984.df = as.data.frame(data1984.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1984.df) <- c("lon", "lat", "Int")
dataInt1985.df = as.data.frame(data1985.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1985.df) <- c("lon", "lat", "Int")
dataInt1986.df = as.data.frame(data1986.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1986.df) <- c("lon", "lat", "Int")
dataInt1987.df = as.data.frame(data1987.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1987.df) <- c("lon", "lat", "Int")
dataInt1988.df = as.data.frame(data1988.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1988.df) <- c("lon", "lat", "Int")
dataInt1989.df = as.data.frame(data1989.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1989.df) <- c("lon", "lat", "Int")
dataInt1990.df = as.data.frame(data1990.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1990.df) <- c("lon", "lat", "Int")
dataInt1991.df = as.data.frame(data1991.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1991.df) <- c("lon", "lat", "Int")
dataInt1992.df = as.data.frame(data1992.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1992.df) <- c("lon", "lat", "Int")
dataInt1993.df = as.data.frame(data1993.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1993.df) <- c("lon", "lat", "Int")
dataInt1994.df = as.data.frame(data1994.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1994.df) <- c("lon", "lat", "Int")
dataInt1995.df = as.data.frame(data1995.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1995.df) <- c("lon", "lat", "Int")
dataInt1996.df = as.data.frame(data1996.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1996.df) <- c("lon", "lat", "Int")
dataInt1997.df = as.data.frame(data1997.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1997.df) <- c("lon", "lat", "Int")
dataInt1998.df = as.data.frame(data1998.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1998.df) <- c("lon", "lat", "Int")
dataInt1999.df = as.data.frame(data1999.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt1999.df) <- c("lon", "lat", "Int")
dataInt2000.df = as.data.frame(data2000.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt2000.df) <- c("lon", "lat", "Int")
dataInt2001.df = as.data.frame(data2001.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt2001.df) <- c("lon", "lat", "Int")
dataInt2002.df = as.data.frame(data2002.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt2002.df) <- c("lon", "lat", "Int")
dataInt2003.df = as.data.frame(data2003.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt2003.df) <- c("lon", "lat", "Int")
dataInt2004.df = as.data.frame(data2004.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt2004.df) <- c("lon", "lat", "Int")
dataInt2005.df = as.data.frame(data2005.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt2005.df) <- c("lon", "lat", "Int")
dataInt2006.df = as.data.frame(data2006.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt2006.df) <- c("lon", "lat", "Int")
dataInt2007.df = as.data.frame(data2007.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt2007.df) <- c("lon", "lat", "Int")
dataInt2008.df = as.data.frame(data2008.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt2008.df) <- c("lon", "lat", "Int")
dataInt2009.df = as.data.frame(data2009.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt2009.df) <- c("lon", "lat", "Int")
dataInt2010.df = as.data.frame(data2010.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt2010.df) <- c("lon", "lat", "Int")
dataInt2011.df = as.data.frame(data2011.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt2011.df) <- c("lon", "lat", "Int")
dataInt2012.df = as.data.frame(data2012.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt2012.df) <- c("lon", "lat", "Int")
dataInt2013.df = as.data.frame(data2013.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt2013.df) <- c("lon", "lat", "Int")
dataInt2014.df = as.data.frame(data2014.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataInt2014.df) <- c("lon", "lat", "Int")

dataInt1975.df$Year[1:50] = 1975
dataInt1976.df$Year[1:50] = 1976
dataInt1977.df$Year[1:50] = 1977
dataInt1978.df$Year[1:50] = 1978
dataInt1979.df$Year[1:50] = 1979
dataInt1980.df$Year[1:50] = 1980
dataInt1981.df$Year[1:50] = 1981
dataInt1982.df$Year[1:50] = 1982
dataInt1983.df$Year[1:50] = 1983
dataInt1984.df$Year[1:50] = 1984
dataInt1985.df$Year[1:50] = 1985
dataInt1986.df$Year[1:50] = 1986
dataInt1987.df$Year[1:50] = 1987
dataInt1988.df$Year[1:50] = 1988
dataInt1989.df$Year[1:50] = 1989
dataInt1990.df$Year[1:50] = 1990
dataInt1991.df$Year[1:50] = 1991
dataInt1992.df$Year[1:50] = 1992
dataInt1993.df$Year[1:50] = 1993
dataInt1994.df$Year[1:50] = 1994
dataInt1995.df$Year[1:50] = 1995
dataInt1996.df$Year[1:50] = 1996
dataInt1997.df$Year[1:50] = 1997
dataInt1998.df$Year[1:50] = 1998
dataInt1999.df$Year[1:50] = 1999
dataInt2000.df$Year[1:50] = 2000
dataInt2001.df$Year[1:50] = 2001
dataInt2002.df$Year[1:50] = 2002
dataInt2003.df$Year[1:50] = 2003
dataInt2004.df$Year[1:50] = 2004
dataInt2005.df$Year[1:50] = 2005
dataInt2006.df$Year[1:50] = 2006
dataInt2007.df$Year[1:50] = 2007
dataInt2008.df$Year[1:50] = 2008
dataInt2009.df$Year[1:50] = 2009
dataInt2010.df$Year[1:50] = 2010
dataInt2011.df$Year[1:50] = 2011
dataInt2012.df$Year[1:50] = 2012
dataInt2013.df$Year[1:50] = 2013
dataInt2014.df$Year[1:50] = 2014

dataInt1975.df$ID = as.numeric(as.factor(with(dataInt1975.df, paste(lon, lat,  sep="_"))))
dataInt1976.df$ID = as.numeric(as.factor(with(dataInt1976.df, paste(lon, lat,  sep="_"))))
dataInt1977.df$ID = as.numeric(as.factor(with(dataInt1977.df, paste(lon, lat,  sep="_"))))
dataInt1978.df$ID = as.numeric(as.factor(with(dataInt1978.df, paste(lon, lat,  sep="_"))))
dataInt1979.df$ID = as.numeric(as.factor(with(dataInt1979.df, paste(lon, lat,  sep="_"))))
dataInt1980.df$ID = as.numeric(as.factor(with(dataInt1980.df, paste(lon, lat,  sep="_"))))
dataInt1981.df$ID = as.numeric(as.factor(with(dataInt1981.df, paste(lon, lat,  sep="_"))))
dataInt1982.df$ID = as.numeric(as.factor(with(dataInt1982.df, paste(lon, lat,  sep="_"))))
dataInt1983.df$ID = as.numeric(as.factor(with(dataInt1983.df, paste(lon, lat,  sep="_"))))
dataInt1984.df$ID = as.numeric(as.factor(with(dataInt1984.df, paste(lon, lat,  sep="_"))))
dataInt1985.df$ID = as.numeric(as.factor(with(dataInt1985.df, paste(lon, lat,  sep="_"))))
dataInt1986.df$ID = as.numeric(as.factor(with(dataInt1986.df, paste(lon, lat,  sep="_"))))
dataInt1987.df$ID = as.numeric(as.factor(with(dataInt1987.df, paste(lon, lat,  sep="_"))))
dataInt1988.df$ID = as.numeric(as.factor(with(dataInt1988.df, paste(lon, lat,  sep="_"))))
dataInt1989.df$ID = as.numeric(as.factor(with(dataInt1989.df, paste(lon, lat,  sep="_"))))
dataInt1990.df$ID = as.numeric(as.factor(with(dataInt1990.df, paste(lon, lat,  sep="_"))))
dataInt1991.df$ID = as.numeric(as.factor(with(dataInt1991.df, paste(lon, lat,  sep="_"))))
dataInt1992.df$ID = as.numeric(as.factor(with(dataInt1992.df, paste(lon, lat,  sep="_"))))
dataInt1993.df$ID = as.numeric(as.factor(with(dataInt1993.df, paste(lon, lat,  sep="_"))))
dataInt1994.df$ID = as.numeric(as.factor(with(dataInt1994.df, paste(lon, lat,  sep="_"))))
dataInt1995.df$ID = as.numeric(as.factor(with(dataInt1995.df, paste(lon, lat,  sep="_"))))
dataInt1996.df$ID = as.numeric(as.factor(with(dataInt1996.df, paste(lon, lat,  sep="_"))))
dataInt1997.df$ID = as.numeric(as.factor(with(dataInt1997.df, paste(lon, lat,  sep="_"))))
dataInt1998.df$ID = as.numeric(as.factor(with(dataInt1998.df, paste(lon, lat,  sep="_"))))
dataInt1999.df$ID = as.numeric(as.factor(with(dataInt1999.df, paste(lon, lat,  sep="_"))))
dataInt2000.df$ID = as.numeric(as.factor(with(dataInt2000.df, paste(lon, lat,  sep="_"))))
dataInt2001.df$ID = as.numeric(as.factor(with(dataInt2001.df, paste(lon, lat,  sep="_"))))
dataInt2002.df$ID = as.numeric(as.factor(with(dataInt2002.df, paste(lon, lat,  sep="_"))))
dataInt2003.df$ID = as.numeric(as.factor(with(dataInt2003.df, paste(lon, lat,  sep="_"))))
dataInt2004.df$ID = as.numeric(as.factor(with(dataInt2004.df, paste(lon, lat,  sep="_"))))
dataInt2005.df$ID = as.numeric(as.factor(with(dataInt2005.df, paste(lon, lat,  sep="_"))))
dataInt2006.df$ID = as.numeric(as.factor(with(dataInt2006.df, paste(lon, lat,  sep="_"))))
dataInt2007.df$ID = as.numeric(as.factor(with(dataInt2007.df, paste(lon, lat,  sep="_"))))
dataInt2008.df$ID = as.numeric(as.factor(with(dataInt2008.df, paste(lon, lat,  sep="_"))))
dataInt2009.df$ID = as.numeric(as.factor(with(dataInt2009.df, paste(lon, lat,  sep="_"))))
dataInt2010.df$ID = as.numeric(as.factor(with(dataInt2010.df, paste(lon, lat,  sep="_"))))
dataInt2011.df$ID = as.numeric(as.factor(with(dataInt2011.df, paste(lon, lat,  sep="_"))))
dataInt2012.df$ID = as.numeric(as.factor(with(dataInt2012.df, paste(lon, lat,  sep="_"))))
dataInt2013.df$ID = as.numeric(as.factor(with(dataInt2013.df, paste(lon, lat,  sep="_"))))
dataInt2014.df$ID = as.numeric(as.factor(with(dataInt2014.df, paste(lon, lat,  sep="_"))))

```

```{r, echo = FALSE}

Intensity = rbind(dataInt1975.df,dataInt1976.df, dataInt1977.df, dataInt1978.df, dataInt1979.df,
                        dataInt1980.df, dataInt1981.df, dataInt1982.df, dataInt1983.df,
                        dataInt1984.df, dataInt1985.df, dataInt1986.df, dataInt1987.df,
                        dataInt1988.df, dataInt1989.df, dataInt1990.df, dataInt1991.df, 
                        dataInt1992.df, dataInt1993.df, dataInt1994.df, dataInt1995.df, 
                        dataInt1996.df, dataInt1997.df, dataInt1998.df, dataInt1999.df,
                        dataInt2000.df, dataInt2001.df, dataInt2002.df, dataInt2003.df,
                        dataInt2004.df, dataInt2005.df, dataInt2006.df, dataInt2007.df,
                        dataInt2008.df, dataInt2009.df, dataInt2010.df, dataInt2011.df,
                        dataInt2012.df, dataInt2013.df, dataInt2014.df) 

Intensity$ID = as.integer(Intensity$ID)

looper = nrow(Intensity)

while(looper > 0)
  {
  if ((Intensity$ID[looper] == 50) == TRUE | (Intensity$ID[looper] == 46) == TRUE |
        (Intensity$ID[looper] == 45) == TRUE | (Intensity$ID[looper] == 35) == TRUE |
        (Intensity$ID[looper] == 49) == TRUE | (Intensity$ID[looper] == 40) == TRUE|
        (Intensity$ID[looper] == 44) == TRUE | (Intensity$ID[looper] == 3) == TRUE |
        (Intensity$ID[looper] == 4) == TRUE)
    {
      Intensity = Intensity[-looper,]
      }
  looper = looper - 1
  }
```

```{r, echo = FALSE}

ID = dataInt1975.df$ID
Yr = c(1975:2014)

IntData.df = data.frame(ID)

IntData.df = cbind(IntData.df, dataInt1975.df$Int)
IntData.df = cbind(IntData.df, dataInt1976.df$Int)
IntData.df = cbind(IntData.df, dataInt1977.df$Int)
IntData.df = cbind(IntData.df, dataInt1978.df$Int)
IntData.df = cbind(IntData.df, dataInt1979.df$Int)
IntData.df = cbind(IntData.df, dataInt1980.df$Int)
IntData.df = cbind(IntData.df, dataInt1981.df$Int)
IntData.df = cbind(IntData.df, dataInt1982.df$Int)
IntData.df = cbind(IntData.df, dataInt1983.df$Int)
IntData.df = cbind(IntData.df, dataInt1984.df$Int)
IntData.df = cbind(IntData.df, dataInt1985.df$Int)
IntData.df = cbind(IntData.df, dataInt1986.df$Int)
IntData.df = cbind(IntData.df, dataInt1987.df$Int)
IntData.df = cbind(IntData.df, dataInt1988.df$Int)
IntData.df = cbind(IntData.df, dataInt1989.df$Int)
IntData.df = cbind(IntData.df, dataInt1990.df$Int)
IntData.df = cbind(IntData.df, dataInt1991.df$Int)
IntData.df = cbind(IntData.df, dataInt1992.df$Int)
IntData.df = cbind(IntData.df, dataInt1993.df$Int)
IntData.df = cbind(IntData.df, dataInt1994.df$Int)
IntData.df = cbind(IntData.df, dataInt1995.df$Int)
IntData.df = cbind(IntData.df, dataInt1996.df$Int)
IntData.df = cbind(IntData.df, dataInt1997.df$Int)
IntData.df = cbind(IntData.df, dataInt1998.df$Int)
IntData.df = cbind(IntData.df, dataInt1999.df$Int)
IntData.df = cbind(IntData.df, dataInt2000.df$Int)
IntData.df = cbind(IntData.df, dataInt2001.df$Int)
IntData.df = cbind(IntData.df, dataInt2002.df$Int)
IntData.df = cbind(IntData.df, dataInt2003.df$Int)
IntData.df = cbind(IntData.df, dataInt2004.df$Int)
IntData.df = cbind(IntData.df, dataInt2005.df$Int)
IntData.df = cbind(IntData.df, dataInt2006.df$Int)
IntData.df = cbind(IntData.df, dataInt2007.df$Int)
IntData.df = cbind(IntData.df, dataInt2008.df$Int)
IntData.df = cbind(IntData.df, dataInt2009.df$Int)
IntData.df = cbind(IntData.df, dataInt2010.df$Int)
IntData.df = cbind(IntData.df, dataInt2011.df$Int)
IntData.df = cbind(IntData.df, dataInt2012.df$Int)
IntData.df = cbind(IntData.df, dataInt2013.df$Int)
IntData.df = cbind(IntData.df, dataInt2014.df$Int)

x = 1
while (x < 41){
  colnames(IntData.df)[x+1] = Yr[x]
  x = x+1
}

IntData.df$ID = as.integer(IntData.df$ID)

looper = nrow(IntData.df)

while(looper > 0)
  {
  if ((IntData.df$ID[looper] == 50) == TRUE | (IntData.df$ID[looper] == 46) == TRUE |
        (IntData.df$ID[looper] == 45) == TRUE | (IntData.df$ID[looper] == 35) == TRUE |
        (IntData.df$ID[looper] == 49) == TRUE | (IntData.df$ID[looper] == 40) == TRUE|
        (IntData.df$ID[looper] == 44) == TRUE | (IntData.df$ID[looper] == 3) == TRUE |
        (IntData.df$ID[looper] == 4) == TRUE)
    {
      IntData.df = IntData.df[-looper,]
      }
  looper = looper - 1
  }

```

```{r, echo = FALSE}

IntStack = stack(data1975.grid, data1976.grid, data1977.grid, data1978.grid, data1979.grid, data1980.grid,
                    data1981.grid, data1982.grid, data1983.grid, data1984.grid, data1985.grid,
                    data1986.grid, data1987.grid, data1988.grid, data1989.grid, data1990.grid,
                    data1991.grid, data1992.grid, data1993.grid, data1994.grid, data1995.grid,
                    data1996.grid, data1997.grid, data1998.grid, data1999.grid, data2000.grid,
                    data2001.grid, data2002.grid, data2003.grid, data2004.grid, data2005.grid,
                    data2006.grid, data2007.grid, data2008.grid, data2009.grid, data2010.grid,
                    data2011.grid, data2012.grid, data2013.grid, data2014.grid)

```

```{r, echo = FALSE}
#This chunk produces Figure 5 in the article.

cr = brewer.pal(7, "Greens")
cr = cr[-(1)]
rng = seq(30, 84, 9)
breaks = c(30, 39, 48, 57, 66, 75, 84 )

plotting = spplot(IntStack, col.regions=cr, names.attr = names, 
                  at = rng, colorkey = list(space="bottom", at=breaks, labels=list(at=breaks)), labels=paste(rng),
                  par.settings=list(fontsize=list(text=10)),
        sub=expression(paste("Intensity [m ", s^-1, "]")),
              pretty=TRUE)

p2 = plotting + latticeExtra::layer(sp.polygons(bPols))
plot(p2)

```


```{r, echo = FALSE}

load("ncSST20142015.RData")
load("ncSST20002013.RData")  
load("ncSST19901999.RData")
load("ncSST19801989.RData")
load("ncSST19701979.RData")

SST = cbind(ncdataframe19701979[,-(3:62)],
            ncdataframe19801989[, 3:length(ncdataframe19801989)], 
            ncdataframe19901999[, 3:length(ncdataframe19901999)], 
            ncSST20002013[, 3:length(ncSST20002013)],
            sst20142015[,3:length(sst20142015)])

sst.sdf = SST
coordinates(sst.sdf) = c("lon", "lat")
proj4string(sst.sdf) = CRS(ll)

```


```{r, echo = FALSE}

sst2 = data.frame(matrix(nrow = nrow(SST), ncol = 42))
colnames(sst2)[1:2] = c("lat", "lon")
colnames(sst2)[3:42] = c(1975:2014)

rowCount = 1
Year = 1975
counter = 8
colCount = 3

while(rowCount <= nrow(SST)){
  sst2$lon[rowCount] = SST$lon[rowCount]
  sst2$lat[rowCount] = SST$lat[rowCount]
  while(counter < ncol(SST)){
    dummy = subset(SST[rowCount, counter:(counter+4)])
    sst2[rowCount, colCount] = rowMeans(dummy)
    counter = counter + 12
    colCount = colCount + 1
    }
  rowCount = rowCount + 1
  Year= Year + 1
  counter = 8
  colCount = 3
}

```

```{r, echo = FALSE, warning=FALSE, message=FALSE}

sst2$mean = rowMeans(sst2[,4:42])

sst2.sdf = sst2
coordinates(sst2.sdf) = c("lon", "lat")
proj4string(sst2.sdf) = CRS(ll)
funParameter = eval(parse(text = "mean"))
sstTotal.grid = rasterize(sst2.sdf, r, 
                         field = 'mean',
                         fun = funParameter)

```

```{r, echo = FALSE}

#This chunk handles mean SST

#data1975 = filter(SST, Yr == 1975)
time = 3 

data1975 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1975.sdf = data1975
coordinates(data1975.sdf) = c("lon", "lat")
proj4string(data1975.sdf) = CRS(ll)

data1975.grid = rasterize(data1975.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1976 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1976.sdf = data1976
coordinates(data1976.sdf) = c("lon", "lat")
proj4string(data1976.sdf) = CRS(ll)

data1976.grid = rasterize(data1976.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1977 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1977.sdf = data1977
coordinates(data1977.sdf) = c("lon", "lat")
proj4string(data1977.sdf) = CRS(ll)

data1977.grid = rasterize(data1977.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1978 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1978.sdf = data1978
coordinates(data1978.sdf) = c("lon", "lat")
proj4string(data1978.sdf) = CRS(ll)

data1978.grid = rasterize(data1978.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1979 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1979.sdf = data1979
coordinates(data1979.sdf) = c("lon", "lat")
proj4string(data1979.sdf) = CRS(ll)

data1979.grid = rasterize(data1979.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1980 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1980.sdf = data1980
coordinates(data1980.sdf) = c("lon", "lat")
proj4string(data1980.sdf) = CRS(ll)

data1980.grid = rasterize(data1980.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1981 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1981.sdf = data1981
coordinates(data1981.sdf) = c("lon", "lat")
proj4string(data1981.sdf) = CRS(ll)

data1981.grid = rasterize(data1981.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1982 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1982.sdf = data1982
coordinates(data1982.sdf) = c("lon", "lat")
proj4string(data1982.sdf) = CRS(ll)

data1982.grid = rasterize(data1982.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1983 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1983.sdf = data1983
coordinates(data1983.sdf) = c("lon", "lat")
proj4string(data1983.sdf) = CRS(ll)

data1983.grid = rasterize(data1983.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1984 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1984.sdf = data1984
coordinates(data1984.sdf) = c("lon", "lat")
proj4string(data1984.sdf) = CRS(ll)

data1984.grid = rasterize(data1984.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1985 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1985.sdf = data1985
coordinates(data1985.sdf) = c("lon", "lat")
proj4string(data1985.sdf) = CRS(ll)

data1985.grid = rasterize(data1985.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1986 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1986.sdf = data1986
coordinates(data1986.sdf) = c("lon", "lat")
proj4string(data1986.sdf) = CRS(ll)

data1986.grid = rasterize(data1986.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1987 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1987.sdf = data1987
coordinates(data1987.sdf) = c("lon", "lat")
proj4string(data1987.sdf) = CRS(ll)

data1987.grid = rasterize(data1987.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1988 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1988.sdf = data1988
coordinates(data1988.sdf) = c("lon", "lat")
proj4string(data1988.sdf) = CRS(ll)

data1988.grid = rasterize(data1988.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1989 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1989.sdf = data1989
coordinates(data1989.sdf) = c("lon", "lat")
proj4string(data1989.sdf) = CRS(ll)

data1989.grid = rasterize(data1989.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1990 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1990.sdf = data1990
coordinates(data1990.sdf) = c("lon", "lat")
proj4string(data1990.sdf) = CRS(ll)

data1990.grid = rasterize(data1990.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1991 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1991.sdf = data1991
coordinates(data1991.sdf) = c("lon", "lat")
proj4string(data1991.sdf) = CRS(ll)

data1991.grid = rasterize(data1991.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1992 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1992.sdf = data1992
coordinates(data1992.sdf) = c("lon", "lat")
proj4string(data1992.sdf) = CRS(ll)

data1992.grid = rasterize(data1992.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1993 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1993.sdf = data1993
coordinates(data1993.sdf) = c("lon", "lat")
proj4string(data1993.sdf) = CRS(ll)

data1993.grid = rasterize(data1993.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1994 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1994.sdf = data1994
coordinates(data1994.sdf) = c("lon", "lat")
proj4string(data1994.sdf) = CRS(ll)

data1994.grid = rasterize(data1994.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1995 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1995.sdf = data1995
coordinates(data1995.sdf) = c("lon", "lat")
proj4string(data1995.sdf) = CRS(ll)

data1995.grid = rasterize(data1995.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1996 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1996.sdf = data1996
coordinates(data1996.sdf) = c("lon", "lat")
proj4string(data1996.sdf) = CRS(ll)

data1996.grid = rasterize(data1996.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1997 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1997.sdf = data1997
coordinates(data1997.sdf) = c("lon", "lat")
proj4string(data1997.sdf) = CRS(ll)

data1997.grid = rasterize(data1997.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1998 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1998.sdf = data1998
coordinates(data1998.sdf) = c("lon", "lat")
proj4string(data1998.sdf) = CRS(ll)

data1998.grid = rasterize(data1998.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data1999 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data1999.sdf = data1999
coordinates(data1999.sdf) = c("lon", "lat")
proj4string(data1999.sdf) = CRS(ll)

data1999.grid = rasterize(data1999.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data2000 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data2000.sdf = data2000
coordinates(data2000.sdf) = c("lon", "lat")
proj4string(data2000.sdf) = CRS(ll)

data2000.grid = rasterize(data2000.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data2001 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data2001.sdf = data2001
coordinates(data2001.sdf) = c("lon", "lat")
proj4string(data2001.sdf) = CRS(ll)

data2001.grid = rasterize(data2001.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data2002 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data2002.sdf = data2002
coordinates(data2002.sdf) = c("lon", "lat")
proj4string(data2002.sdf) = CRS(ll)

data2002.grid = rasterize(data2002.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data2003 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data2003.sdf = data2003
coordinates(data2003.sdf) = c("lon", "lat")
proj4string(data2003.sdf) = CRS(ll)

data2003.grid = rasterize(data2003.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data2004 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data2004.sdf = data2004
coordinates(data2004.sdf) = c("lon", "lat")
proj4string(data2004.sdf) = CRS(ll)

data2004.grid = rasterize(data2004.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data2005 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data2005.sdf = data2005
coordinates(data2005.sdf) = c("lon", "lat")
proj4string(data2005.sdf) = CRS(ll)

data2005.grid = rasterize(data2005.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data2006 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data2006.sdf = data2006
coordinates(data2006.sdf) = c("lon", "lat")
proj4string(data2006.sdf) = CRS(ll)

data2006.grid = rasterize(data2006.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data2007 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data2007.sdf = data2007
coordinates(data2007.sdf) = c("lon", "lat")
proj4string(data2007.sdf) = CRS(ll)

data2007.grid = rasterize(data2007.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data2008 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data2008.sdf = data2008
coordinates(data2008.sdf) = c("lon", "lat")
proj4string(data2008.sdf) = CRS(ll)

data2008.grid = rasterize(data2008.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data2009 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data2009.sdf = data2009
coordinates(data2009.sdf) = c("lon", "lat")
proj4string(data2009.sdf) = CRS(ll)

data2009.grid = rasterize(data2009.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data2010 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data2010.sdf = data2010
coordinates(data2010.sdf) = c("lon", "lat")
proj4string(data2010.sdf) = CRS(ll)

data2010.grid = rasterize(data2010.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data2011 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data2011.sdf = data2011
coordinates(data2011.sdf) = c("lon", "lat")
proj4string(data2011.sdf) = CRS(ll)

data2011.grid = rasterize(data2011.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data2012 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data2012.sdf = data2012
coordinates(data2012.sdf) = c("lon", "lat")
proj4string(data2012.sdf) = CRS(ll)

data2012.grid = rasterize(data2012.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data2013 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data2013.sdf = data2013
coordinates(data2013.sdf) = c("lon", "lat")
proj4string(data2013.sdf) = CRS(ll)

data2013.grid = rasterize(data2013.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

data2014 = cbind(sst2[,1:2],
                 sst = sst2[,time])
time = time + 1
data2014.sdf = data2014
coordinates(data2014.sdf) = c("lon", "lat")
proj4string(data2014.sdf) = CRS(ll)

data2014.grid = rasterize(data2014.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

datasst1975.df = as.data.frame(data1975.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1975.df) <- c("lon", "lat", "sst")
datasst1976.df = as.data.frame(data1976.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1976.df) <- c("lon", "lat", "sst")
datasst1977.df = as.data.frame(data1977.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1977.df) <- c("lon", "lat", "sst")
datasst1978.df = as.data.frame(data1978.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1978.df) <- c("lon", "lat", "sst")
datasst1979.df = as.data.frame(data1979.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1979.df) <- c("lon", "lat", "sst")
datasst1980.df = as.data.frame(data1980.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1980.df) <- c("lon", "lat", "sst")
datasst1981.df = as.data.frame(data1981.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1981.df) <- c("lon", "lat", "sst")
datasst1982.df = as.data.frame(data1982.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1982.df) <- c("lon", "lat", "sst")
datasst1983.df = as.data.frame(data1983.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1983.df) <- c("lon", "lat", "sst")
datasst1984.df = as.data.frame(data1984.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1984.df) <- c("lon", "lat", "sst")
datasst1985.df = as.data.frame(data1985.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1985.df) <- c("lon", "lat", "sst")
datasst1986.df = as.data.frame(data1986.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1986.df) <- c("lon", "lat", "sst")
datasst1987.df = as.data.frame(data1987.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1987.df) <- c("lon", "lat", "sst")
datasst1988.df = as.data.frame(data1988.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1988.df) <- c("lon", "lat", "sst")
datasst1989.df = as.data.frame(data1989.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1989.df) <- c("lon", "lat", "sst")
datasst1990.df = as.data.frame(data1990.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1990.df) <- c("lon", "lat", "sst")
datasst1991.df = as.data.frame(data1991.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1991.df) <- c("lon", "lat", "sst")
datasst1992.df = as.data.frame(data1992.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1992.df) <- c("lon", "lat", "sst")
datasst1993.df = as.data.frame(data1993.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1993.df) <- c("lon", "lat", "sst")
datasst1994.df = as.data.frame(data1994.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1994.df) <- c("lon", "lat", "sst")
datasst1995.df = as.data.frame(data1995.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1995.df) <- c("lon", "lat", "sst")
datasst1996.df = as.data.frame(data1996.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1996.df) <- c("lon", "lat", "sst")
datasst1997.df = as.data.frame(data1997.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1997.df) <- c("lon", "lat", "sst")
datasst1998.df = as.data.frame(data1998.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1998.df) <- c("lon", "lat", "sst")
datasst1999.df = as.data.frame(data1999.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst1999.df) <- c("lon", "lat", "sst")
datasst2000.df = as.data.frame(data2000.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst2000.df) <- c("lon", "lat", "sst")
datasst2001.df = as.data.frame(data2001.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst2001.df) <- c("lon", "lat", "sst")
datasst2002.df = as.data.frame(data2002.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst2002.df) <- c("lon", "lat", "sst")
datasst2003.df = as.data.frame(data2003.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst2003.df) <- c("lon", "lat", "sst")
datasst2004.df = as.data.frame(data2004.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst2004.df) <- c("lon", "lat", "sst")
datasst2005.df = as.data.frame(data2005.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst2005.df) <- c("lon", "lat", "sst")
datasst2006.df = as.data.frame(data2006.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst2006.df) <- c("lon", "lat", "sst")
datasst2007.df = as.data.frame(data2007.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst2007.df) <- c("lon", "lat", "sst")
datasst2008.df = as.data.frame(data2008.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst2008.df) <- c("lon", "lat", "sst")
datasst2009.df = as.data.frame(data2009.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst2009.df) <- c("lon", "lat", "sst")
datasst2010.df = as.data.frame(data2010.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst2010.df) <- c("lon", "lat", "sst")
datasst2011.df = as.data.frame(data2011.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst2011.df) <- c("lon", "lat", "sst")
datasst2012.df = as.data.frame(data2012.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst2012.df) <- c("lon", "lat", "sst")
datasst2013.df = as.data.frame(data2013.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst2013.df) <- c("lon", "lat", "sst")
datasst2014.df = as.data.frame(data2014.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(datasst2014.df) <- c("lon", "lat", "sst")

datasst1975.df$Year[1:50] = 1975
datasst1976.df$Year[1:50] = 1976
datasst1977.df$Year[1:50] = 1977
datasst1978.df$Year[1:50] = 1978
datasst1979.df$Year[1:50] = 1979
datasst1980.df$Year[1:50] = 1980
datasst1981.df$Year[1:50] = 1981
datasst1982.df$Year[1:50] = 1982
datasst1983.df$Year[1:50] = 1983
datasst1984.df$Year[1:50] = 1984
datasst1985.df$Year[1:50] = 1985
datasst1986.df$Year[1:50] = 1986
datasst1987.df$Year[1:50] = 1987
datasst1988.df$Year[1:50] = 1988
datasst1989.df$Year[1:50] = 1989
datasst1990.df$Year[1:50] = 1990
datasst1991.df$Year[1:50] = 1991
datasst1992.df$Year[1:50] = 1992
datasst1993.df$Year[1:50] = 1993
datasst1994.df$Year[1:50] = 1994
datasst1995.df$Year[1:50] = 1995
datasst1996.df$Year[1:50] = 1996
datasst1997.df$Year[1:50] = 1997
datasst1998.df$Year[1:50] = 1998
datasst1999.df$Year[1:50] = 1999
datasst2000.df$Year[1:50] = 2000
datasst2001.df$Year[1:50] = 2001
datasst2002.df$Year[1:50] = 2002
datasst2003.df$Year[1:50] = 2003
datasst2004.df$Year[1:50] = 2004
datasst2005.df$Year[1:50] = 2005
datasst2006.df$Year[1:50] = 2006
datasst2007.df$Year[1:50] = 2007
datasst2008.df$Year[1:50] = 2008
datasst2009.df$Year[1:50] = 2009
datasst2010.df$Year[1:50] = 2010
datasst2011.df$Year[1:50] = 2011
datasst2012.df$Year[1:50] = 2012
datasst2013.df$Year[1:50] = 2013
datasst2014.df$Year[1:50] = 2014

datasst1975.df$ID = as.numeric(as.factor(with(datasst1975.df, paste(lon, lat,  sep="_"))))
datasst1976.df$ID = as.numeric(as.factor(with(datasst1976.df, paste(lon, lat,  sep="_"))))
datasst1977.df$ID = as.numeric(as.factor(with(datasst1977.df, paste(lon, lat,  sep="_"))))
datasst1978.df$ID = as.numeric(as.factor(with(datasst1978.df, paste(lon, lat,  sep="_"))))
datasst1979.df$ID = as.numeric(as.factor(with(datasst1979.df, paste(lon, lat,  sep="_"))))
datasst1980.df$ID = as.numeric(as.factor(with(datasst1980.df, paste(lon, lat,  sep="_"))))
datasst1981.df$ID = as.numeric(as.factor(with(datasst1981.df, paste(lon, lat,  sep="_"))))
datasst1982.df$ID = as.numeric(as.factor(with(datasst1982.df, paste(lon, lat,  sep="_"))))
datasst1983.df$ID = as.numeric(as.factor(with(datasst1983.df, paste(lon, lat,  sep="_"))))
datasst1984.df$ID = as.numeric(as.factor(with(datasst1984.df, paste(lon, lat,  sep="_"))))
datasst1985.df$ID = as.numeric(as.factor(with(datasst1985.df, paste(lon, lat,  sep="_"))))
datasst1986.df$ID = as.numeric(as.factor(with(datasst1986.df, paste(lon, lat,  sep="_"))))
datasst1987.df$ID = as.numeric(as.factor(with(datasst1987.df, paste(lon, lat,  sep="_"))))
datasst1988.df$ID = as.numeric(as.factor(with(datasst1988.df, paste(lon, lat,  sep="_"))))
datasst1989.df$ID = as.numeric(as.factor(with(datasst1989.df, paste(lon, lat,  sep="_"))))
datasst1990.df$ID = as.numeric(as.factor(with(datasst1990.df, paste(lon, lat,  sep="_"))))
datasst1991.df$ID = as.numeric(as.factor(with(datasst1991.df, paste(lon, lat,  sep="_"))))
datasst1992.df$ID = as.numeric(as.factor(with(datasst1992.df, paste(lon, lat,  sep="_"))))
datasst1993.df$ID = as.numeric(as.factor(with(datasst1993.df, paste(lon, lat,  sep="_"))))
datasst1994.df$ID = as.numeric(as.factor(with(datasst1994.df, paste(lon, lat,  sep="_"))))
datasst1995.df$ID = as.numeric(as.factor(with(datasst1995.df, paste(lon, lat,  sep="_"))))
datasst1996.df$ID = as.numeric(as.factor(with(datasst1996.df, paste(lon, lat,  sep="_"))))
datasst1997.df$ID = as.numeric(as.factor(with(datasst1997.df, paste(lon, lat,  sep="_"))))
datasst1998.df$ID = as.numeric(as.factor(with(datasst1998.df, paste(lon, lat,  sep="_"))))
datasst1999.df$ID = as.numeric(as.factor(with(datasst1999.df, paste(lon, lat,  sep="_"))))
datasst2000.df$ID = as.numeric(as.factor(with(datasst2000.df, paste(lon, lat,  sep="_"))))
datasst2001.df$ID = as.numeric(as.factor(with(datasst2001.df, paste(lon, lat,  sep="_"))))
datasst2002.df$ID = as.numeric(as.factor(with(datasst2002.df, paste(lon, lat,  sep="_"))))
datasst2003.df$ID = as.numeric(as.factor(with(datasst2003.df, paste(lon, lat,  sep="_"))))
datasst2004.df$ID = as.numeric(as.factor(with(datasst2004.df, paste(lon, lat,  sep="_"))))
datasst2005.df$ID = as.numeric(as.factor(with(datasst2005.df, paste(lon, lat,  sep="_"))))
datasst2006.df$ID = as.numeric(as.factor(with(datasst2006.df, paste(lon, lat,  sep="_"))))
datasst2007.df$ID = as.numeric(as.factor(with(datasst2007.df, paste(lon, lat,  sep="_"))))
datasst2008.df$ID = as.numeric(as.factor(with(datasst2008.df, paste(lon, lat,  sep="_"))))
datasst2009.df$ID = as.numeric(as.factor(with(datasst2009.df, paste(lon, lat,  sep="_"))))
datasst2010.df$ID = as.numeric(as.factor(with(datasst2010.df, paste(lon, lat,  sep="_"))))
datasst2011.df$ID = as.numeric(as.factor(with(datasst2011.df, paste(lon, lat,  sep="_"))))
datasst2012.df$ID = as.numeric(as.factor(with(datasst2012.df, paste(lon, lat,  sep="_"))))
datasst2013.df$ID = as.numeric(as.factor(with(datasst2013.df, paste(lon, lat,  sep="_"))))
datasst2014.df$ID = as.numeric(as.factor(with(datasst2014.df, paste(lon, lat,  sep="_"))))

```

```{r, echo = FALSE}

SSTdata = rbind(datasst1975.df,datasst1976.df, datasst1977.df, datasst1978.df, datasst1979.df,
                        datasst1980.df, datasst1981.df, datasst1982.df, datasst1983.df,
                        datasst1984.df, datasst1985.df, datasst1986.df, datasst1987.df,
                        datasst1988.df, datasst1989.df, datasst1990.df, datasst1991.df, 
                        datasst1992.df, datasst1993.df, datasst1994.df, datasst1995.df, 
                        datasst1996.df, datasst1997.df, datasst1998.df, datasst1999.df,
                        datasst2000.df, datasst2001.df, datasst2002.df, datasst2003.df,
                        datasst2004.df, datasst2005.df, datasst2006.df, datasst2007.df,
                        datasst2008.df, datasst2009.df, datasst2010.df, datasst2011.df,
                        datasst2012.df, datasst2013.df, datasst2014.df) 

SSTdata$ID = as.integer(SSTdata$ID)

looper = nrow(SSTdata)

while(looper > 0)
  {
  if ((SSTdata$ID[looper] == 50) == TRUE | (SSTdata$ID[looper] == 46) == TRUE |
        (SSTdata$ID[looper] == 45) == TRUE | (SSTdata$ID[looper] == 35) == TRUE |
        (SSTdata$ID[looper] == 49) == TRUE | (SSTdata$ID[looper] == 40) == TRUE|
        (SSTdata$ID[looper] == 44) == TRUE | (SSTdata$ID[looper] == 3) == TRUE |
        (SSTdata$ID[looper] == 4) == TRUE)
    {
      SSTdata = SSTdata[-looper,]
      }
  looper = looper - 1
  }
```

```{r, echo = FALSE}
 
ID = datasst1975.df$ID

sstData.df = data.frame(ID)

sstData.df = cbind(sstData.df, datasst1975.df$sst)
sstData.df = cbind(sstData.df, datasst1976.df$sst)
sstData.df = cbind(sstData.df, datasst1977.df$sst)
sstData.df = cbind(sstData.df, datasst1978.df$sst)
sstData.df = cbind(sstData.df, datasst1979.df$sst)
sstData.df = cbind(sstData.df, datasst1980.df$sst)
sstData.df = cbind(sstData.df, datasst1981.df$sst)
sstData.df = cbind(sstData.df, datasst1982.df$sst)
sstData.df = cbind(sstData.df, datasst1983.df$sst)
sstData.df = cbind(sstData.df, datasst1984.df$sst)
sstData.df = cbind(sstData.df, datasst1985.df$sst)
sstData.df = cbind(sstData.df, datasst1986.df$sst)
sstData.df = cbind(sstData.df, datasst1987.df$sst)
sstData.df = cbind(sstData.df, datasst1988.df$sst)
sstData.df = cbind(sstData.df, datasst1989.df$sst)
sstData.df = cbind(sstData.df, datasst1990.df$sst)
sstData.df = cbind(sstData.df, datasst1991.df$sst)
sstData.df = cbind(sstData.df, datasst1992.df$sst)
sstData.df = cbind(sstData.df, datasst1993.df$sst)
sstData.df = cbind(sstData.df, datasst1994.df$sst)
sstData.df = cbind(sstData.df, datasst1995.df$sst)
sstData.df = cbind(sstData.df, datasst1996.df$sst)
sstData.df = cbind(sstData.df, datasst1997.df$sst)
sstData.df = cbind(sstData.df, datasst1998.df$sst)
sstData.df = cbind(sstData.df, datasst1999.df$sst)
sstData.df = cbind(sstData.df, datasst2000.df$sst)
sstData.df = cbind(sstData.df, datasst2001.df$sst)
sstData.df = cbind(sstData.df, datasst2002.df$sst)
sstData.df = cbind(sstData.df, datasst2003.df$sst)
sstData.df = cbind(sstData.df, datasst2004.df$sst)
sstData.df = cbind(sstData.df, datasst2005.df$sst)
sstData.df = cbind(sstData.df, datasst2006.df$sst)
sstData.df = cbind(sstData.df, datasst2007.df$sst)
sstData.df = cbind(sstData.df, datasst2008.df$sst)
sstData.df = cbind(sstData.df, datasst2009.df$sst)
sstData.df = cbind(sstData.df, datasst2010.df$sst)
sstData.df = cbind(sstData.df, datasst2011.df$sst)
sstData.df = cbind(sstData.df, datasst2012.df$sst)
sstData.df = cbind(sstData.df, datasst2013.df$sst)
sstData.df = cbind(sstData.df, datasst2014.df$sst)

x = 1
while (x < 41){
  colnames(sstData.df)[x+1] = Yr[x]
  x = x+1
}

sstData.df$ID = as.integer(sstData.df$ID)

looper = nrow(sstData.df)

while(looper > 0)
  {
  if ((sstData.df$ID[looper] == 50) == TRUE | (sstData.df$ID[looper] == 46) == TRUE |
        (sstData.df$ID[looper] == 45) == TRUE | (sstData.df$ID[looper] == 35) == TRUE |
        (sstData.df$ID[looper] == 49) == TRUE | (sstData.df$ID[looper] == 40) == TRUE|
        (sstData.df$ID[looper] == 44) == TRUE | (sstData.df$ID[looper] == 3) == TRUE |
        (sstData.df$ID[looper] == 4) == TRUE)
    {
      sstData.df = sstData.df[-looper,]
      }
  looper = looper - 1
}


```

```{r, echo = FALSE}

sstStack = stack(data1975.grid, data1976.grid, data1977.grid, data1978.grid, data1979.grid, data1980.grid,
                    data1981.grid, data1982.grid, data1983.grid, data1984.grid, data1985.grid,
                    data1986.grid, data1987.grid, data1988.grid, data1989.grid, data1990.grid,
                    data1991.grid, data1992.grid, data1993.grid, data1994.grid, data1995.grid,
                    data1996.grid, data1997.grid, data1998.grid, data1999.grid, data2000.grid,
                    data2001.grid, data2002.grid, data2003.grid, data2004.grid, data2005.grid,
                    data2006.grid, data2007.grid, data2008.grid, data2009.grid, data2010.grid,
                    data2011.grid, data2012.grid, data2013.grid, data2014.grid)

```

```{r, echo = FALSE}

cr = brewer.pal(7, "Reds")
cr = cr[-(1:2)]
#cr = cr[-length(cr)]
rng = seq(20, 30, 2)
breaks = c(20, 22, 24, 26, 28, 30)

plotting = spplot(sstStack, col.regions=cr, names.attr = names, 
                  at = rng, colorkey = list(space="bottom", at=breaks, labels=list(at=breaks)), labels=paste(rng),
                  par.settings=list(fontsize=list(text=10)),
      sub=expression(paste("Sea surface temperature [", degree, "C]")),
              pretty=TRUE)

p3 = plotting + latticeExtra::layer(sp.polygons(bPols))
#p3

```

```{r, echo = FALSE}

dataSSTanom.df = as.data.frame(sstTotal.grid, row.names=NULL, optional = TRUE, xy = TRUE, centroids = TRUE)
colnames(dataSSTanom.df) <- c("lon", "lat", "sst")

dataSSTanom.df$sst[1:13] = NA

dataSSTanom.sdf = dataSSTanom.df
coordinates(dataSSTanom.sdf) = c("lon", "lat")
proj4string(dataSSTanom.sdf) = CRS(ll)

dataSSTanom.grid = rasterize(dataSSTanom.sdf, r, 
                         field = 'sst',
                         fun = funParameter)

```

```{r, echo = FALSE}
#This chunk produces Figure 6 in the article.

cr1 = rev(brewer.pal(5, "Blues"))
cr1 = cr1[-c(4:5)]
cr2 = brewer.pal(5, "Reds")
cr2 = cr2[-c(1:2)]
cr = c(cr1,cr2)
#cr = cr[-length(cr)]
rng = seq(-1.5, 1.5, 0.5)
breaks = c(-1.5, -1.0, -0.5, 0.0, 0.5, 1.0, 1.5)

plotting = spplot(sstStack - dataSSTanom.grid, col.regions=cr, names.attr = names, 
                  at = rng, colorkey = list(space="bottom", at=breaks, labels=list(at=breaks)), labels=paste(rng),
                  par.settings=list(fontsize=list(text=10)),
      sub=expression(paste("Sea surface temperature anomalies [", degree, "C]")),
              pretty=TRUE)

p3a = plotting + latticeExtra::layer(sp.polygons(bPols))
plot(p3a)

```


```{r, echo = FALSE}

data.df = merge(Intensification, Intensity)
data.df = merge(data.df, SSTdata)

load("mjoRMM.RData")

mjoRMM$rmmSum = mjoRMM$meanRMM1 + mjoRMM$meanRMM2

data.df["RMM1"] = NA
data.df["RMM2"] = NA

data.df  = data.df[with(data.df, order(Year)),]

Yr = 1975
counter = 1
loop = 1

while (Yr < 2014)
  {
   while ((data.df$Year[counter] == Yr)== TRUE){  
      data.df$RMM1[counter] = mjoRMM$meanRMM1[loop]
      data.df$RMM2[counter] = mjoRMM$meanRMM2[loop]
       data.df$rmmSum[counter] = mjoRMM$rmmSum[loop]
       counter = counter + 1
      }
    Yr = Yr + 1
    loop = loop + 1
  }
#The error that occurs when running the above loop does not
#affect any of the data or code. Therefore, it can be ignored.

```

```{r, echo = FALSE}

mei = read.table("meiForMJO.txt", header=TRUE)
#mei = mei[-5,] 
mei = mei[-1,]
mei = mei[,-(2:6)]
mei = mei[,-(7:8)]

mei$avg = rowMeans(subset(mei, select = c(Jun, Jul, Aug, Sep, Oct)))

Yr = 1975
counter = 1
loop = 2

while (Yr < 2014)
  {
    while ((data.df$Year[counter] == Yr)== TRUE){  
      data.df$mei[counter] = mei$avg[loop]
      counter = counter + 1
      }
    Yr = Yr + 1
    loop = loop + 1
}
#The error that occurs when running the above loop does not
#affect any of the data or code. Therefore, it can be ignored.

```

```{r, echo = FALSE}

nao = read.table("naoData.txt", header=TRUE)


nao = nao[,-(5:13)]

nao$avg = rowMeans(subset(nao, select = c(Jan, Feb, Mar)))

Yr = 1975
counter = 1
loop = 2

while (Yr < 2014)
  {
    while ((data.df$Year[counter] == Yr)== TRUE){  
      data.df$nao[counter] = nao$avg[loop]
      
      counter = counter + 1
      }
    Yr = Yr + 1
    loop = loop + 1
}

naoYears = nao[order(nao$avg),]
negNAO = naoYears$year[1:10]
posNAO = naoYears$year[31:40]

#The error that occurs when running the above loop does not
#affect any of the data or code. Therefore, it can be ignored.
```
Figure 7. Time series of (a) monthly average of MEI from 1975 to 2014, (b) monthly average of NAOI from 1975 to 2014...
```{r, echo = FALSE}
#This chunk produces Figure 7 in the article.
library(grid)
library(gridExtra)
library(plotrix)
library(reshape2)

meiTime = ggplot(data=mei, aes(x=Yr,y=avg, group=1)) +
  geom_line(size=1.25) + ylab("MEI [s.d.]") + xlab("") +
  theme(plot.title=element_text(hjust=0)) +
  ggtitle('(a)')

naoTime = ggplot(data=nao, aes(x=year,y=avg, group=1)) +
  geom_line(size=1.25) + ylab("NAOI [s.d.]") + xlab("") +
  theme(plot.title=element_text(hjust=0)) +
  ggtitle('(b)')

mjoTime = ggplot(data=mjoRMM, aes(x=year,y=rmmSum, group=1)) +
  geom_line(size=1.25) + ylab("MJOI [s.d.]") + xlab("Year") +
  theme(plot.title=element_text(hjust=0)) +
  ggtitle('(c)')

grid.arrange(meiTime, naoTime, mjoTime, widths=unit(0.5,"npc"), ncol=1)

```

```{r, echo = FALSE}
#This chunk produces Figure 3b in the article.

mean(Tracks.grid@data@values, na.rm=TRUE)

range(values(Tracks.grid), na.rm=TRUE)
cr = brewer.pal(6, "Purples")
cr = cr[-(1:2)]
rng = seq(0, 0.80, 0.20)
breaks = c(0.00, 0.20, 0.40, 0.60, 0.80)
vals = levelplot(Tracks.grid, margin = FALSE,
          sub=expression(paste("Intensification [m ", s^-1, " per h]")),
          xlab = NULL, ylab = NULL, 
          at=rng, col.regions = cr,
          colorkey = list(space="bottom", at=breaks, labels=list(at=breaks)), labels=paste(rng),
          border = "white", border.lwd = 2,
          par.settings = list(fontsize=list(text=12)),
          scales=list(draw=FALSE))
pDIntDt = vals + latticeExtra::layer(sp.polygons(bPols, col = gray(.8))) 
pDIntDt = update(pDIntDt, main = textGrob("b", x = unit(.05, "npc"), 
                               gp = gpar(fontsize=16)))
plot(pDIntDt)

```
Figure 11. Density plot of five strongest mean intensification values (scaled to 24 h) for each of the 10 highest NAOI years (pink)...
```{r, echo = FALSE}
#This chunk produces Figure 11 in the article.

DIntDtmaxNAO = data.frame(Y2002 = dataDIntDt2002.df$DIntDt,
                          Y2014 = dataDIntDt2014.df$DIntDt,
                          Y1997 = dataDIntDt1997.df$DIntDt,
                          Y1989 = dataDIntDt1989.df$DIntDt,
                          Y1994 = dataDIntDt1994.df$DIntDt,
                          Y1993 = dataDIntDt1993.df$DIntDt,
                          Y2012 = dataDIntDt2012.df$DIntDt,
                          Y2000 = dataDIntDt2000.df$DIntDt,
                          Y1995 = dataDIntDt1995.df$DIntDt,
                          Y1990 = dataDIntDt1990.df$DIntDt)

DIntDtminNAO = data.frame(Y2010 = dataDIntDt2010.df$DIntDt,
                          Y1977 = dataDIntDt1977.df$DIntDt,
                          Y1985 = dataDIntDt1985.df$DIntDt,
                          Y1987 = dataDIntDt1987.df$DIntDt,
                          Y2013 = dataDIntDt2013.df$DIntDt,
                          Y1979 = dataDIntDt1979.df$DIntDt,
                          Y1980 = dataDIntDt1980.df$DIntDt,
                          Y1978 = dataDIntDt1978.df$DIntDt,
                          Y1975 = dataDIntDt1975.df$DIntDt,
                          Y2001 = dataDIntDt2001.df$DIntDt)

minNAOdata = data.frame(lapply(DIntDtminNAO, sort, decreasing=TRUE, na.last=TRUE))
maxNAOdata = data.frame(lapply(DIntDtmaxNAO, sort, decreasing=TRUE, na.last=TRUE))

minNAODIntDt = as.vector(as.matrix(minNAOdata[1:5,]))
maxNAODIntDt = as.vector(as.matrix(maxNAOdata[1:5,]))
minmax = 24
naoData = data.frame(maxNAO = maxNAODIntDt, minNAO = minNAODIntDt)
naoData$minNAO = naoData$minNAO * minmax #max is here; its 43.8
naoData$maxNAO = naoData$maxNAO * minmax

mNAOdata = melt(naoData)

plot(ggplot(mNAOdata, aes(value, fill=variable)) +
  geom_density(alpha=.3, color=NA) +
  scale_fill_discrete(name=NULL, 
                      breaks=c("maxNAO", "minNAO"),
                      labels=c("10 Highest NAOI averaged years",
        "10 Lowest NAOI averaged years")) +
  xlab("Intensification [m"~s^{-1}~"per day]") +
  ylab("Density") + theme(legend.position = c(.73,.9)) +
  theme(legend.title=element_blank(),
        axis.text.x = element_text(size=20),
        axis.text.y = element_text(size=20),
        axis.title.x = element_text(size=20),
        axis.title.y = element_text(size=20),
        legend.text = element_text(size=16)) +
  geom_linerange(x=15.43, ymax=0.08, ymin=0, color="white", size=2))

```




Figure 3. Mean intensification histogram and spatial plot. (a) Histogram of mean hourly intensification...
```{r, echo = FALSE}
#This chunk produces Figure 3a in the article, as well as the entire Figure 3.
PlotFigure3a = function(data.df){
histDIntDt = ggplot(data.df, aes(DIntDt)) + geom_histogram(binwidth=0.2,
  color = "white", fill = "#6A51A3") +
  xlab(expression(paste("Mean Intensification [m ", s^-1, " per h]"))) +
  ylab("Number of Grid Cells") 
histDIntDt = histDIntDt + ggtitle("a") + theme(plot.title = element_text(hjust = 0))
plot(histDIntDt)
grid.arrange(histDIntDt, pDIntDt,widths=unit(0.5,"npc"), ncol=1)
}
PlotFigure3a(data.df)
```

```{r, echo = FALSE, message=FALSE}
years = c(1975:2014)

Cov = data.frame(Year = mei$Yr, mei = mei$avg, nao = nao$avg, 
                 rmmSum = mjoRMM$rmmSum)

#install.packages("INLA", repos=c(getOption("repos"), INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)
require(INLA)
require(spdep)
test.nb = poly2nb(test)
nb2INLA("testnb.INLA", test.nb)
test.nb = inla.read.graph("testnb.INLA")

st = expand.grid(ID = 1:nrow(DIntDtData.df), Year = years) 
data.INLA = merge(st, Cov, by="Year")
DIntDtData.df = DIntDtData.df[,-1]
data.INLA$DIntDt = as.vector(as.matrix(DIntDtData.df)) 
IntData.df = IntData.df[,-1]
data.INLA$Int = as.vector(as.matrix(IntData.df)) 
sstData.df = sstData.df[,-1]
data.INLA$SST = as.vector(as.matrix(sstData.df))
#ohcData.df = ohcData.df[,-1]
#data.INLA$OHC = as.vector(as.matrix(ohcData.df))
data.INLA$ID.1 = data.INLA$ID.2 = data.INLA$ID.3 = 
  data.INLA$ID.4 = data.INLA$ID.5 =
  data.INLA$ID

```

```{r, echo = FALSE}
model="besag"
formula = DIntDt ~ Int + SST + mei + rmmSum + nao +
  f(ID.1, Int, model=model, graph=test.nb) +
  f(ID.2, SST, model=model, graph=test.nb) +
  f(ID.3, mei, model=model, graph=test.nb) + 
  f(ID.4, rmmSum, model=model, graph=test.nb) + 
  f(ID.5, nao, model=model, graph=test.nb)

```


```{r, echo = FALSE}

model.inla = inla(formula, family="gamma", data = data.INLA,
                  control.compute=list(dic=TRUE)) 
summary(model.inla)

```




```{r, echo = FALSE}
#This chunk produces Figure 2a in the article. 
PlotFigure2a = function(Tracks.sdf, r){
library(RColorBrewer)
library(wesanderson)
Count.grid = rasterize(Tracks.sdf, r,
                        field = 'Sid',
                        fun = function(x,...) {length(unique(na.omit(x)))})
range(values(Count.grid), na.rm=TRUE)
rng = seq(0, 35, 5)
crq = wes_palette(7, name="Rushmore", type="continuous")
vals = levelplot(Count.grid, margin = FALSE,
          sub=expression(paste("Number of Intensifying Hurricanes per cell")),
          xlab = NULL, ylab = NULL, 
          at=rng, col.regions = crq,
          colorkey = list(space='bottom'),
          border = "white", border.lwd = 2,
          par.settings = list(fontsize=list(text=12)),
          scales=list(draw=FALSE)) 
pCount = vals + latticeExtra::layer(sp.polygons(bPols, col = gray(.8))) 
pCount = update(pCount, main = textGrob("a", x = unit(.05, "npc"), 
                               gp = gpar(fontsize=16)))
plot(pCount)
}
PlotFigure2a(Tracks.sdf,r)
```
Figure 2. Hurricane intensification count and ratio per cell. (a) Number of intensifying hurricanes per cell...
```{r, echo = FALSE}
#This chunk produces Figure 2b in the article, as well as the entire Figure 2.

PlotFigure2b = function(fullData.sdf,r){
fullCount.grid = rasterize(fullData.sdf, r,
                        field = 'Sid',
                        fun = function(x,...) {length(unique(na.omit(x)))})

difCount.grid = Count.grid / fullCount.grid

rng = seq(0.2, 1, 0.2)
breaks =  c(0.2, 0.4, 0.6, 0.8, 1.0)
cr = brewer.pal(5, "Blues")
cr = cr[-(1)]
vals = levelplot(difCount.grid, margin = FALSE,
          sub=expression(paste("Ratio of intensifying hurricanes to all hurricanes")),
          xlab = NULL, ylab = NULL, 
          at=rng, col.regions = cr,
          colorkey = list(space = 'bottom',    at=breaks,labels=list(at=breaks)),
          border = "white", border.lwd = 2,
          par.settings = list(fontsize=list(text=12)),
          scales=list(draw=FALSE))
pDif = vals + latticeExtra::layer(sp.polygons(bPols, col = gray(.8))) 
pDif = update(pDif, main = textGrob("b", x = unit(.05, "npc"), 
                               gp = gpar(fontsize=16)))
plot(pDif)
grid.arrange(pCount, pDif, ncol=1)
}
PlotFigure2b(fullData.sdf,r)
```

```{r, echo = FALSE}
#This chunk produces Figures 8 and 10 in the article.

intPost = as.data.frame(model.inla$marginals.fixed$Int)

intPostPlot = ggplot(data=intPost, aes(x=x,y=y, group=1)) +
  geom_line() + xlab(expression(paste("",beta[Int]," [m ",s^-1,h^-1," per m ",s^-1,"]"))) + ylab("Density") +
  theme(axis.text.x = element_text(size=20),
  axis.text.y = element_text(size=20), text = element_text(size=20),
  plot.title=element_text(hjust=0)) +
  geom_linerange(x=0, ymin=0, ymax=80, color="red", size=3) +
  geom_linerange(x=-0.0003, ymin=0, ymax = 80, color="gray", size=3)+
  geom_linerange(x=0.0213, ymin=0, ymax = 80, color="gray", size=3)+
  ggtitle('a')
plot(intPostPlot)
sstPost = as.data.frame(model.inla$marginals.fixed$SST)

sstPostPlot = ggplot(data=sstPost, aes(x=x,y=y, group=1)) +
  geom_line() + xlab(expression(paste("",beta[SST]," [m ", s^-1, h^-1," per ",degree,"C]"))) + ylab("Density") +
  theme(axis.text.x = element_text(size=20),
  axis.text.y = element_text(size=20), text = element_text(size=20),
  plot.title=element_text(hjust=0)) +
  geom_linerange(x=0, ymin=0, ymax=30, color="red", size=3) +
  geom_linerange(x=0.0525, ymin=0, ymax = 30, color="gray", size=3)+
  geom_linerange(x=0.1102, ymin=0, ymax = 30, color="gray", size=3)+
  ggtitle('b')
plot(sstPostPlot)
#grid.arrange(intPostPlot, sstPostPlot, heights=c(1,1)))

meiPost = as.data.frame(model.inla$marginals.fixed$mei)

meiPostPlot = ggplot(data=meiPost, aes(x=x,y=y, group=1)) +
  geom_line() + xlab(expression(paste("",beta[MEI]," [m ", s^-1, h^-1," per 1 SD]"))) + ylab("Density") +
  theme(axis.text.x = element_text(size=20),
  axis.text.y = element_text(size=20), text = element_text(size=20),
  plot.title=element_text(hjust=0)) +
  geom_linerange(x=0, ymin=0, ymax=8, color="red", size=2) +
  geom_linerange(x=-0.1045, ymin=0, ymax = 8, color="gray", size=2)+
  geom_linerange(x=0.1155, ymin=0, ymax = 8, color="gray", size=2)+
  ggtitle('a')
plot(meiPostPlot)
naoPost = as.data.frame(model.inla$marginals.fixed$nao)

naoPostPlot = ggplot(data=naoPost, aes(x=x,y=y, group=1)) +
  geom_line() + xlab(expression(paste("",beta[NAOI]," [m ", s^-1, h^-1," per 1 SD]"))) + ylab("Density") +
  theme(axis.text.x = element_text(size=20),
  axis.text.y = element_text(size=20), text = element_text(size=20),
  plot.title=element_text(hjust=0)) +
  geom_linerange(x=0, ymin=0, ymax=7, color="red", size=2) +
  geom_linerange(x=-0.3004, ymin=0, ymax = 7, color="gray", size=2)+
  geom_linerange(x=-0.0514, ymin=0, ymax = 7, color="gray", size=2)+
  ggtitle('b')
plot(naoPostPlot)
mjoPost = as.data.frame(model.inla$marginals.fixed$rmmSum)

mjoPostPlot = ggplot(data=mjoPost, aes(x=x,y=y, group=1)) +
  geom_line() + xlab(expression(paste("",beta[MJOI]," [m ", s^-1, h^-1," per 1 SD]"))) + ylab("Density") +
  theme(axis.text.x = element_text(size=20),
  axis.text.y = element_text(size=20), text = element_text(size=20),
  plot.title=element_text(hjust=0)) +
  geom_linerange(x=0, ymin=0, ymax=2.5, color="red", size=2) +
  geom_linerange(x=-0.1974, ymin=0, ymax = 2.5, color="gray", size=2)+
  geom_linerange(x=0.5443, ymin=0, ymax = 2.5, color="gray", size=2)+
  ggtitle('c')
plot(mjoPostPlot)
#plot(grid.arrange(meiPostPlot, naoPostPlot, mjoPostPlot, heights=c(1,1,1)))

```

```{r, echo = FALSE}
coef.Int = model.inla$summary.random[[1]]$mean + model.inla$summary.fixed["Int","mean"]
coef.Int.sd = model.inla$summary.random[[1]]$sd + model.inla$summary.fixed["Int","sd"]

coef.SST = model.inla$summary.random[[2]]$mean + model.inla$summary.fixed["SST","mean"]
coef.SST.sd = model.inla$summary.random[[2]]$sd + model.inla$summary.fixed["SST","sd"]

#coef.error = model.inla$summary.random[[6]]$mean + model.inla$summary.fixed["(Intercept)", "mean"] 
#coef.error.sd = model.inla$summary.random[[6]]$sd + model.inla$summary.fixed["(Intercept)", "sd"] 

coef.cov.df = data.frame(coef.Int, coef.Int.sd, coef.SST, coef.SST.sd)

#cor.test(coef.Int, coef.error) #0.991 -> spatially confounded
#cor.test(coef.SST, coef.error) #0.966 -> spatially confounded

```

```{r, echo = FALSE}

coef.cov.df$lat = latOrder 
coef.cov.df$lon = lonOrder 

FinalData.sdf = coef.cov.df
coordinates(FinalData.sdf) = c("lon", "lat")
proj4string(FinalData.sdf) = CRS(ll)

```

```{r, echo = FALSE}
#This chunk produces Figure 9a in the article.

Int.grid = rasterize(FinalData.sdf, r, 
                         field = 'coef.Int',
                         fun = mean)

range(values(Int.grid), na.rm = TRUE)
rng = seq(0.003, 0.017, .002)
breaks = c(0.003, 0.005, 0.007, 0.009, 0.011, 0.013, 0.015, 0.017)
cr = brewer.pal(9, "Greens")
cr = cr[-(1:2)]
vals = levelplot(Int.grid, margin = FALSE, 
          xlab = NULL, ylab = NULL, 
          col.regions = cr, at = rng, 
          colorkey = list(space = 'bottom', at=breaks, labels=list(at=breaks)),
          border = "white", border.lwd = 2,
          par.settings = list(fontsize = list(text = 20)),
          scales=list(draw=FALSE),
          sub=expression(paste("",bar(beta)[Int],"(s) [m ",s^-1,h^-1," per m ",s^-1,"]")))
pInt = vals + latticeExtra::layer(sp.polygons(bPols, col = gray(.8)))
pInt = update(pInt, main = textGrob("a", x = unit(.05, "npc"), 
                               gp = gpar(fontsize=16)))
plot(pInt)

```

```{r, echo = FALSE}
#This chunk produces Figure 9b in the article, as well as the entire Figure 9.

library(gridExtra)

SST.grid = rasterize(FinalData.sdf, r, 
                         field = 'coef.SST',
                         fun = mean)

mean(SST.grid@data@values, na.rm=TRUE)

range(values(SST.grid), na.rm = TRUE)
rng = seq(0.077, 0.085, .001)
breaks = c(0.077, 0.078, 0.079, 0.080, 0.081, 0.082, 0.083, 0.084, 0.085)
cr = brewer.pal(9, "Reds")
cr = cr[-(1)]
vals = levelplot(SST.grid, margin = FALSE, 
          xlab = NULL, ylab = NULL, 
          col.regions = cr, at = rng, 
          colorkey = list(space = 'bottom',    at=breaks,labels=list(at=breaks)),
          border = "white", border.lwd = 2,
          par.settings = list(fontsize = list(text = 20)),
          scales=list(draw=FALSE),
          sub=expression(paste("",bar(beta)[SST],"(s) [m ",s^-1,h^-1," per ",degree,"C ]")))
pSST = vals + latticeExtra::layer(sp.polygons(bPols, col = gray(.8)))
pSST = update(pSST, main = textGrob("b", x = unit(.05, "npc"), 
                               gp = gpar(fontsize=16)))
plot(pSST)

grid.arrange(pInt, pSST, ncol=1)

```
